---
title: "QuantAQ Missing Data Report"
author: "Lewis White"
date: "2023-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### load packages 
```{r}
library(QuantAQAPIClient)
library(here) 
library(lubridate) 
library(tictoc)
library(DT)
library(purrr)
library(tidyverse)
```

### Connect to QuantAQ API (necessary to obtain data)
```{r}
setup_client() #log in to QuantAQ account and click developer in the left menu to grab API key
```

```{r}
Sys.setenv(TZ='UTC') #SET TIME ZONE

Sys.setenv(TZ = 'Africa/Kampala') #EAT for kenya, uganda, ethiopia 
```

### source functions
```{r}
#function to obtain data by date for list of devices

#version 1
source(here("get_data_for_devices.R"))

#version 2
source(here("get_full_data_test.R"))
```

### create lists of devices by country 

```{r}
uganda_devices <- c("MOD-PM-00844", "MOD-PM-00847", "MOD-PM-00831", "MOD-PM-00846", "MOD-PM-00832",
                     "MOD-PM-00845", "MOD-PM-00828", "MOD-PM-00826", "MOD-PM-00838", "MOD-00117",
                     "MOD-PM-00836")

kenya_devices <- c("MOD-PM-00852", "MOD-PM-00834", "MOD-PM-00830", "MOD-PM-00840", "MOD-PM-00843",
                    "MOD-PM-00841", "MOD-PM-00849", "MOD-PM-00850", "MOD-PM-00848", "MOD-PM-00829",
                    "MOD-00116")

ethiopia_devices <- c("MOD-PM-00827", "MOD-PM-00839", "MOD-PM-00842", "MOD-PM-00833", "MOD-PM-00855",
                    "MOD-PM-00854", "MOD-PM-00835", "MOD-PM-00837", "MOD-PM-00851", "MOD-00118",
                    "MOD-PM-00853")

ghana_devices <- c("MOD-PM-01054", "MOD-PM-00900", "MOD-PM-00876", "MOD-PM-00882", "MOD-PM-00896",
                    "MOD-PM-00897", "MOD-PM-00892", "MOD-PM-00877", "MOD-PM-01060", "MOD-PM-01055",
                    "MOD-PM-00884", "MOD-PM-01056", "MOD-PM-01051", "MOD-PM-01059", "MOD-PM-00881",
                    "MOD-PM-00891", "MOD-PM-00898", "MOD-PM-01052", "MOD-00400", "MOD-PM-00894",
                    "MOD-PM-01053", "MOD-PM-00887", "MOD-PM-00886", "MOD-PM-00879", "MOD-PM-00890",
                    "MOD-PM-00889", "MOD-PM-00899", "MOD-PM-00883", "MOD-PM-00895", "MOD-PM-01057",
                    "MOD-PM-01058", "MOD-PM-00893", "MOD-PM-00878", "MOD-PM-00888", "MOD-PM-00885",
                    "MOD-00398", "MOD-00401", "MOD-00399", "MOD-00397", "MOD-PM-00880")
```


## Using get_data_for_devices.R
```{r}
# start_date <- as.Date("2023-10-15")
# end_date <- as.Date("2023-11-15")
# device_names <- c("MOD-PM-00826", "MOD-PM-00847")
# 
# result_df <- get_data_for_devices(start_date, end_date, device_names)
# 
# result_df_2 <- get_full_data_test(start_date, end_date, device_names)
# 
# # Calculate missing hours based on the absence of timestamp
# result_df_2 %>%
#   group_by(date, hour, monitor) %>%
#   summarise(no_data = sum(is.na(pm25))) %>%
#   mutate(missing_hour = ifelse(no_data > 30, 1, 0)) %>%
#   ungroup() %>%
#   group_by(date, monitor) %>%
#   summarise(hours_missing_day = sum(missing_hour)) %>%
#   pivot_wider(names_from = date, values_from = hours_missing_day) %>%
#   datatable()

```



```{r}
# Define start_date and end_date
start_date <- as.Date("2023-11-15")
end_date <- as.Date("2023-12-01")

# List of device serial numbers
device_list <- uganda_devices

# Initialize an empty list to store results for each device
result_combined <- list()

# Loop through each device in the device list
for (device in device_list) {
  # Function to get data by date, handling errors
  get_data_safe <- possibly(get_data_by_date, otherwise = NULL)
  
  # Use map to get data for each date, handling errors
  result_list <- map(seq(start_date, end_date, by = "days"), function(date) {
    formatted_date <- format(date, "%Y-%m-%d")
    get_data_safe(sn = device, date = formatted_date)
  })
  
  # Filter out NULL elements (empty lists)
  result_list <- purrr::discard(result_list, ~ is.null(.x) || length(.x) == 0)
  
  if (!is_empty(result_list)) {
  # Combine the list of data frames into a single data frame
  result_df <- do.call(rbind, lapply(result_list, as.data.frame)) %>%
    mutate(monitor = device) %>%
    select(monitor, everything()) %>%
    mutate(timestamp = as.POSIXct(timestamp)) %>% 
    mutate(timestamp = format(timestamp, "%Y-%m-%d %H:%M")) %>% 
    mutate(timestamp = lubridate::ymd_hm(timestamp))
  
  minutely_df <- data.frame(timestamp = seq.POSIXt(
    as.POSIXct(start_date, tz = "UTC"),
    as.POSIXct(end_date + 1, tz = "UTC"),
    by = "min"
  ))
  
  result_df_full <- full_join(result_df, minutely_df) %>% 
    arrange(timestamp) %>%
    mutate(date = as.Date(timestamp)) %>%  
    mutate(hour = hour(ymd_hms(timestamp))) %>%
    mutate(monitor = device) %>%
    select(monitor, timestamp, date, hour, everything()) 
  
  # Store the result for the current device in the combined list
  result_combined[[device]] <- result_df_full
  } else {
    # If there's no data for this device, create an empty dataframe
    minutely_df_empty <- data.frame(timestamp = seq.POSIXt(
      as.POSIXct(start_date, tz = "UTC"),
      as.POSIXct(end_date + 1, tz = "UTC"),
      by = "min"
    ))
    
    minutely_df_empty <- minutely_df_empty %>% mutate(date = as.Date(timestamp)) %>%  
      mutate(hour = hour(ymd_hms(timestamp))) %>%
      mutate(monitor = device) %>%
      select(monitor, timestamp, date, hour) 
      
    result_combined[[device]] <- minutely_df_empty
  }
}

# Combine data for all devices into a single data frame
final_result_df <- bind_rows(result_combined)

#check with very active device
uganda11 <- final_result_df %>% filter(monitor == "MOD-00117")

#check with less active device 
uganda1 <- final_result_df %>% filter(monitor == "MOD-PM-00826")

final_result_df %>%
  filter(date <= as.Date(end_date)) %>%
  group_by(date, hour, monitor) %>%
  summarise(no_data = sum(is.na(pm25))) %>%
  mutate(missing_hour = ifelse(no_data > 30, 1, 0)) %>%
  ungroup() %>%
  group_by(date, monitor) %>%
  summarise(hours_missing_day = sum(missing_hour)) %>%
  pivot_wider(names_from = date, values_from = hours_missing_day) %>%
  datatable()
```

First time is local 3am. So it is 3 hours earlier than what the QuantAQ monitors will say. 

The get_data_by_date() uses timestamp instead of timestamp_local

```{r}
#add 3 hours to timestamp so it aligns 
uganda_shifted_time <- uganda11 %>%
  mutate(timestamp = ymd_hms(timestamp)) %>%
  mutate(timestamp = timestamp + hours(3))
```





## START DATE NEEDS TO BE 1 DAY BEFORE THE ACTUAL START DATE OF INTEREST

```{r}
# Define start_date and end_date
start_date <- as.Date("2023-11-14")
end_date <- as.Date("2023-12-01")

# List of device serial numbers
device_list <- uganda_devices

# Initialize an empty list to store results for each device
result_combined <- list()

# Loop through each device in the device list
for (device in device_list) {
  # Function to get data by date, handling errors
  get_data_safe <- possibly(get_data_by_date, otherwise = NULL)
  
  # Use map to get data for each date, handling errors
  result_list <- map(seq(start_date, end_date, by = "days"), function(date) {
    formatted_date <- format(date, "%Y-%m-%d")
    get_data_safe(sn = device, date = formatted_date)
  })
  
  # Filter out NULL elements (empty lists)
  result_list <- purrr::discard(result_list, ~ is.null(.x) || length(.x) == 0)
  
  if (!is_empty(result_list)) {
  # Combine the list of data frames into a single data frame
  result_df <- do.call(rbind, lapply(result_list, as.data.frame)) %>%
    mutate(monitor = device) %>%
    select(monitor, everything()) %>%
    mutate(timestamp = as.POSIXct(timestamp)) %>% 
    mutate(timestamp = format(timestamp, "%Y-%m-%d %H:%M")) %>% 
    mutate(timestamp = lubridate::ymd_hm(timestamp)) %>%
    mutate(local_timestamp = timestamp + hours(3))
  
  minutely_df <- data.frame(timestamp = seq.POSIXt(
    as.POSIXct(start_date, tz = "UTC"),
    as.POSIXct(end_date + 1, tz = "UTC"),
    by = "min"
  )) %>%
    mutate(local_timestamp = timestamp + hours(3))
  
  result_df_full <- full_join(result_df, minutely_df) %>% 
    arrange(timestamp) %>%
    mutate(date = as.Date(local_timestamp)) %>%  
    mutate(hour = hour(ymd_hms(local_timestamp))) %>%
    mutate(monitor = device) %>%
    select(monitor, timestamp, local_timestamp, date, hour, everything()) 
  
  # Store the result for the current device in the combined list
  result_combined[[device]] <- result_df_full
  } else {
    # If there's no data for this device, create an empty dataframe
    minutely_df_empty <- data.frame(timestamp = seq.POSIXt(
      as.POSIXct(start_date, tz = "UTC"),
      as.POSIXct(end_date + 1, tz = "UTC"),
      by = "min"
    )) %>%
    mutate(local_timestamp = timestamp + hours(3))
    
    minutely_df_empty <- minutely_df_empty %>% mutate(date = as.Date(local_timestamp)) %>%  
      mutate(hour = hour(ymd_hms(local_timestamp))) %>%
      mutate(monitor = device) %>%
      select(monitor, timestamp, local_timestamp, date, hour) 
      
    result_combined[[device]] <- minutely_df_empty
  }
}

# Combine data for all devices into a single data frame
final_result_df <- bind_rows(result_combined)

#check with very active devices
uganda11 <- final_result_df %>% filter(monitor == "MOD-00117")

#check with less active device 
uganda1 <- final_result_df %>% filter(monitor == "MOD-PM-00826")

final_result_df %>%
  filter(date <= as.Date(end_date),
         date > as.Date(start_date)) %>%
  group_by(date, hour, monitor) %>%
  summarise(no_data = sum(is.na(pm25))) %>%
  mutate(missing_hour = ifelse(no_data > 30, 1, 0)) %>%
  ungroup() %>%
  group_by(date, monitor) %>%
  summarise(hours_missing_day = sum(missing_hour)) %>%
  pivot_wider(names_from = date, values_from = hours_missing_day) %>%
  datatable()
```

First time is local 3am. So it is 3 hours earlier than what the QuantAQ monitors will say. 

The get_data_by_date() uses timestamp instead of timestamp_local





















## Using get_full_data_test.R

```{r}
# # In order for the code to run, each device needs to have been online for at least one observation between the specified start date and end date. Otherwise, the following error will result: 
# 
# # "Error in `map()`:
# # ℹ In index: 9.
# # Caused by error in `UseMethod()`:
# # ! no applicable method for 'mutate' applied to an object of class "NULL""
# 
# 
# start_date <- as.Date("2023-11-25")
# end_date <- as.Date("2023-12-05")
# device_names <- c(
#   "MOD-PM-00826", "MOD-PM-00847", "MOD-PM-00846", "MOD-PM-00832", 
#   "MOD-PM-00844", "MOD-PM-00845", "MOD-PM-00838", "MOD-PM-00828", 
#   "MOD-PM-00836", "MOD-PM-00831", "MOD-00117")
# 
# device_names <- c(
#   "MOD-PM-00826", "MOD-PM-00847", "MOD-PM-00845")
# 
# result <- get_full_data_test(start_date, end_date, device_names)
# 
# result %>%
#   group_by(date, hour, monitor) %>%
#   summarise(no_data = sum(is.na(pm25))) %>%
#   mutate(missing_hour = ifelse(no_data > 30, 1, 0)) %>%
#   ungroup() %>%
#   group_by(date, monitor) %>%
#   summarise(hours_missing_day = sum(missing_hour)) %>%
#   pivot_wider(names_from = date, values_from = hours_missing_day) %>%
#   datatable()
# 
# Kenya7 <- result %>% filter(monitor == "MOD-PM-00843")
```




## troubleshooting case where device is offline for total period of interest

```{r}
# # Wrap the function `get_data_by_date` within `possibly` to handle potential errors
# get_data_safe <- possibly(get_data_by_date, otherwise = NULL)
# 
# 
# get_full_data_test <- function(start_date, end_date, device_names) {
#   start_date <- as.POSIXct(start_date)
#   end_date <- as.POSIXct(end_date)
#   
#   result_df <- map_df(device_names, function(device) {
#     result_list <- map(seq(start_date, end_date, by = "days"), function(date) {
#       formatted_date <- format(date, "%Y-%m-%d")
#       get_data_safe(sn = device, date = formatted_date)
#     }) %>%
#       discard(~ is.null(.x) || length(.x) == 0) %>%
#       map(as.data.frame)
#     
#     result_df_device <- do.call(rbind, result_list)
#     
#     if (!is.null(result_df_device)) {
#       result_df_device <- result_df_device %>%
#         mutate(
#           timestamp = as.POSIXct(timestamp),
#           timestamp = format(timestamp, "%Y-%m-%d %H:%M"),
#           timestamp2 = as.POSIXct(timestamp),
#           date = as.Date(timestamp2),
#           hour = hour(ymd_hms(timestamp2)),
#           monitor = device
#         ) %>%
#         complete(timestamp2 = seq.POSIXt(start_date, end_date, by = "min")) %>%
#         select(monitor, everything())
#     } else {
#       result_df_device <- data.frame()  # Providing an empty dataframe as the else condition
#     }
#     
#     return(result_df_device)
#   })
#   
#   return(result_df)
# }

```


```{r}
# start_date <- as.Date("2023-11-15")
# end_date <- as.Date("2023-12-01")
# 
# 
# test_devices <- c("MOD-PM-00838", "MOD-00117",
#                      "MOD-PM-00836")
# 
# result <- get_full_data_test(start_date, end_date, device_names = uganda_devices)

### ISSUE: It isn't recording the monitor for missing data points!
```




