---
title: "Ghana Gas Report"
output: 
  html_document:
    toc: true
    theme: united
date: "2024-06-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
### load packages 
library(here) # file path org
library(lubridate)# working with dates
library(tictoc) # timing
library(DT) # datatables
library(purrr) # applying functions across df
library(tidyverse) # data cleaning and plotting
library(data.table) 
library(sf) # spatial data 
library(viridis) # color pallete 
library(knitr)
library(modelsummary) # table of regressions
library(spdep)
library(gstat)
library(units) 
library(gridExtra)
library(broom)
library(Metrics) 
library(kableExtra) # table creation

```


```{r source functions, echo=FALSE, message=FALSE, warning=FALSE}
# Source in functions

# Source function that creates fleet average vs monitor pollutant reading plot
source(here("src", "compare_fleet_avg_monitor.R"))

# Source function that creates heatmap of pollutant readings
source(here("src", "generate_heatmap.R"))

# Source function to make map of pollution values at each monitor location
source(here("src", "generate_spatial_pollution_map.R"))

# source function to calculate moran i 
source(here("src", "calculate_moran_i.R"))
```

```{r location data, echo=FALSE, message=FALSE, warning=FALSE}
#LOAD LOCATION DATA ----
monitor_community_info <- read_csv(here("data", "monitor_community_info.csv"))

# create spatial feature dataset of monitor points
#monitor_points <- st_as_sf(monitor_community_info, coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))


monitor_points <- monitor_community_info %>%
  st_as_sf(coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))


# LOAD IN THE COUNTRY AND REGION MAP DATA ----

country <- st_read(here("data", "spatial", "ghana_boundaries_shp", "gha_admbnda_adm0_gss_20210308.shp"), quiet = TRUE)

regions <- st_read(here("data", "spatial", "ghana_boundaries_shp",  "gha_admbnda_adm1_gss_20210308.shp"), quiet = TRUE) %>%
  rename(region = ADM1_EN)

bono_east <- regions %>% filter(region == "Bono East")


## SPECIFIC TOWN MARKERS ----

# just the lat/lon for kintampo
kintampo_sf <- st_as_sf(data.frame(
  location = "Kintampo",
  geo_lon = -1.7296,
  geo_lat = 8.0593
), coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))

# Dwenewoho
dwenewoho_sf <- st_as_sf(data.frame(
  location = "dwenewoho",
  geo_lon = -1.8325,
  geo_lat = 7.7403
), coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))


# New Longoro
new_longoro_sf <- st_as_sf(data.frame(
  location = "New Longoro",
  geo_lon = -2.02953,
  geo_lat = 8.14399
), coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))

# Apesika
apesika_sf <- st_as_sf(data.frame(
  location = "Apesika",
  geo_lon = -1.5453,
  geo_lat = 7.99838
), coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))

# Kurawura Akura
kurawura_akura_sf <- st_as_sf(data.frame(
  location = "Kurawura Akura",
  geo_lon = -1.4684,
  geo_lat = 8.7380
), coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))


# LOAD IN THE ROADS DATA ----

# Load roads data and set CRS (assuming it's EPSG:4326)
roads <- st_read(here("data", "spatial", "ghana_roads_shp", "hotosm_gha_roads_lines_shp.shp"), quiet = TRUE)
st_crs(roads) <- 4326

#Filter the roads to only bono east
roads_filtered <- st_intersection(roads, bono_east)
```



## Monitor Locations

```{r plot monitor locations, echo=FALSE, message=FALSE, warning=FALSE}
# Define the bounding box coordinates
bbox_coords <- matrix(c(-2.2, 7.6, -2.2, 8.8, -1.3, 8.8, -1.3, 7.6, -2.2, 7.6), ncol = 2, byrow = TRUE)
bbox_polygon <- st_polygon(list(bbox_coords))
bbox_sf <- st_sfc(bbox_polygon, crs = st_crs(4326))

# Plot the map with the bounding box
ggplot() +
  # Add Ghana regions
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5) +
  # Add monitor points with transparency
  geom_sf(data = monitor_points, aes(geometry = geometry), color = "red", alpha = 0.5, size = 2) +
  # Add bounding box
  geom_sf(data = bbox_sf, fill = NA, color = "blue", lwd = 0.5, linetype = "solid") +
  # Add Ghana country boundary
  geom_sf(data = country, fill = NA, color = "black", size = 1) +
  #labs(title = "Monitor Locations in Ghana") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 1), # Increase tick mark size
    plot.title = element_text(size = 16) 
  )


ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5) +
  # Add monitor points with transparency within the bounding box
  geom_sf(data = monitor_points, aes(geometry = geometry, color = monitor_type), alpha = 0.5, size = 5) +
  
  # Add Kintampo marker
  geom_sf(data = kintampo_sf, aes(geometry = geometry), shape = 20, size = 3, color = "black") +
  # Add Dwenewoho marker
  geom_sf(data = dwenewoho_sf, aes(geometry = geometry), shape = 20, size = 3, color = "black") +
  # Add new longoro marker
  geom_sf(data = new_longoro_sf, aes(geometry = geometry), shape = 20, size = 3, color = "black") +
  # Add apesika marker
  geom_sf(data = apesika_sf, aes(geometry = geometry), shape = 20, size = 3, color = "black") +
  # Add kurawura akura marker
  geom_sf(data = kurawura_akura_sf, aes(geometry = geometry), shape = 20, size = 3, color = "black") +
  
  # Add Kintamp annotation text
  annotate("text", x = -1.73, y = 8.10, label = "Kintampo", color = "black", size = 5) +
  # Add Dwenewoho annotation text
  annotate("text", x = -1.69, y = 7.74, label = "Dwenewoho", color = "black", size = 5) +
  # Add new longoro annotation text
  annotate("text", x = -2.047, y = 8.214, label = "New Longoro", color = "black", size = 5) +
  # Add apesika annotation text
  annotate("text", x = -1.459, y = 7.96, label = "Apesika", color = "black", size = 5) +
  # Add atta akura annotation text
  annotate("text", x = -1.65, y = 8.75, label = "Kurawura Akura", color = "black", size = 5) +
  
  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  labs(x = "", 
       y = "",
       color = "Monitor Type") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.5), # Increase tick mark size
    plot.title = element_text(size = 16), # Increase title size
    legend.text = element_text(size = 10), # Increase legend text size
    legend.title = element_text(size = 12) # Increase legend title size
  )
```


## Load corrected pollution data and create summaries 

```{r}
## LOAD GAS DATA ----

# load full gas data
gas_corrected_full <- readRDS(here("data", "gas", "final", "corrected_community_gas_20230926-20240816.rds"))

# load individual gas data summarized
co_community_hourly <- readRDS(here("data", "gas", "summarized", "co_community_hourly_20231024-20240816.rds"))
co_community_daily <- readRDS(here("data", "gas", "summarized", "co_community_daily_20231024-20240816.rds"))

no_community_hourly <- readRDS(here("data", "gas", "summarized", "no_community_hourly_20231024-20240816.rds"))
no_community_daily <- readRDS(here("data", "gas", "summarized", "no_community_daily_20231024-20240816.rds"))

no2_community_hourly <- readRDS(here("data", "gas", "summarized", "no2_community_hourly_20231024-20240816.rds"))
no2_community_daily <- readRDS(here("data", "gas", "summarized", "no2_community_daily_20231024-20240816.rds"))

o3_community_hourly <- readRDS(here("data", "gas", "summarized", "o3_community_hourly_20231024-20240816.rds"))
o3_community_daily <- readRDS(here("data", "gas", "summarized", "o3_community_daily_20231024-20240816.rds"))
```



```{r}
#monitor missingness
monitor_missingness <- gas_corrected_full %>% 
  group_by(monitor) %>% 
  summarise(missing_rate = mean(is.na(o3)) * 100) %>%
  mutate(monitor = fct_reorder(monitor, -missing_rate))

# MAP OF MONITOR MISSINGNESS BY LOCATION ----
missing_monitor_location <- left_join(monitor_missingness, monitor_points)
```


## Data source (Cloud vs SD for each monitor)

```{r plot cloud vs sd data, echo=FALSE, message=FALSE, warning=FALSE}
monitor_obs_count_type <- gas_corrected_full %>%
  group_by(monitor, source) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  group_by(monitor) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ungroup() %>%
  filter(!is.na(source)) %>%
  mutate(source = factor(source, levels = c("sd_card", "cloud")))

monitor_order <- monitor_obs_count_type %>%
  filter(source %in% c("cloud", "sd_card")) %>%
  group_by(monitor) %>%
  summarise(total = sum(percentage)) %>%
  arrange(desc(total)) %>%
  pull(monitor)

monitor_obs_count_type <- monitor_obs_count_type %>%
  mutate(monitor = factor(monitor, levels = monitor_order))

ggplot(monitor_obs_count_type, aes(x = monitor, y = percentage, fill = source)) +
  geom_bar(stat = "identity") +
  labs(title = "",
       x = "Monitor",
       y = "Percentage of Data Recorded",
       fill = "Data Source") +  # Custom legend title
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_text(size = 14),  # Increase x-axis title size
        axis.title.y = element_text(size = 14),  # Increase y-axis title size
        legend.text = element_text(size = 12),   # Increase legend text size
        legend.title = element_text(size = 14)) +  # Increase legend title size
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  scale_fill_manual(values = c("cloud" = "#2599db", "sd_card" = "#b54033"),
                    labels = c("cloud" = "Cloud", "sd_card" = "SD Card"))


```



# Gas Pollution Trends

## Grand Average plots 

```{r grand average facet, echo=FALSE, message=FALSE, warning=FALSE}
# join daily PM into one dataset
daily_pollutants <- left_join(co_community_daily, no_community_daily) %>% 
  left_join(no2_community_daily) %>%
  left_join(o3_community_daily)


daily_pollutants$date <- as.Date(daily_pollutants$date)


# pivot longer data so each measurement is its own row 
daily_pollutants_long <- daily_pollutants %>%
  group_by(date) %>%
  summarise('Mean CO' = mean(mean_co, na.rm = TRUE),
            'Mean NO' = mean(mean_no, na.rm = TRUE),
            'Mean NO2' = mean(mean_no2, na.rm = TRUE),
            'Mean O3' = mean(mean_o3, na.rm = TRUE)) %>%
  ungroup() %>%
  pivot_longer(cols = c("Mean CO", "Mean NO", "Mean NO2", "Mean O3"), names_to = "pollutant", values_to = "measurement")

# Calculate the mean values for each pollutant
mean_values <- daily_pollutants_long %>%
  group_by(pollutant) %>%
  summarise(mean_measurement = signif(mean(measurement, na.rm = TRUE), 3))

# Merge the mean values back into the original data frame for plotting
daily_pollutants_long <- daily_pollutants_long %>%
  left_join(mean_values, by = "pollutant")


# Convert the 'pollutant' column to a factor with the desired order
daily_pollutants_long$pollutant <- factor(daily_pollutants_long$pollutant, levels = c("Mean CO", "Mean NO", "Mean NO2", "Mean O3"))


daily_pollutants_long %>%
  ggplot(aes(x = date, y = measurement)) +
  geom_line() +
  geom_hline(aes(yintercept = mean_measurement), color = "#cc1414", linetype = "dashed") +
  geom_text(data = mean_values, aes(x = max(daily_pollutants_long$date), y = mean_measurement, label = paste0(pollutant, ": ", round(mean_measurement, 2), " ppb")),
            color = "#cc1414", hjust = 1, vjust = -4.5, size = 4.5) +
  theme_bw() +
  labs(x = "Date",
       y = "Gas Measurement (ppb)") +
  facet_grid(factor(pollutant, levels = c("Mean CO", "Mean NO", "Mean NO2", "Mean O3")) ~ ., scales = "free_y") +
  theme(
    axis.title = element_text(size = 14),   # Increase axis titles size
    axis.text = element_text(size = 12),    # Increase axis labels size
    strip.text = element_text(size = 12)  # Increase facet group text size
  )

```



# Carbon Monixide Community Deployment

### Checking Individual Monitor to Fleet Average

```{r monitor vs fleet avg 25, fig.height=10, fig.width=7, echo=FALSE, message=FALSE, warning=FALSE}
# hourly comparison
compare_fleet_avg_monitor(co_community_hourly, "co", "hourly")

# Daily comparison
compare_fleet_avg_monitor(co_community_daily, "co", "daily")
```

### PM Carbon Monixide Trends

```{r co timelines, echo=FALSE, message=FALSE, warning=FALSE}
daily_co_all_devices <- gas_corrected_full %>%
  group_by(date, monitor) %>%
  summarize(mean_co = mean(co, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_co, group = monitor)) +
  geom_line(color = "black", alpha = 0.2) +
  theme_minimal() +
  labs(y = "Mean CO",
       title = "Average CO Accross All Devices")

daily_cooktime_co_all_devices <- gas_corrected_full %>%
  filter(hour >= 16 & hour < 20) %>%
  group_by(monitor, date) %>%
  summarise(mean_co = mean(co, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_co, group = monitor)) +
  geom_line(color = "black", alpha = 0.4) +
  theme_minimal() +
  labs(y = "Mean CO",
       title = "Daily Average CO for Each Monitor During Primary Cooking Hours (4-8)") 

grid.arrange(daily_co_all_devices, daily_cooktime_co_all_devices)

daily_co_timeline <- gas_corrected_full %>%
  group_by(date) %>%
  summarize(mean_co = mean(co, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_co)) +
  geom_line() +
  theme_minimal() +
  labs(y = "Mean PM CO",
       title = "Average CO Accross All Devices")

daily_cooktime_co_timeline <- gas_corrected_full %>%
  filter(hour >= 16 & hour < 20) %>%
  group_by(date) %>%
  summarize(mean_co = mean(co, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_co)) +
  geom_line() +
  theme_minimal() +
  labs(y = "Mean PM CO",
       title = "Mean CO Accross All Devices During Cooking Hours (4-8)")

grid.arrange(daily_co_timeline, daily_cooktime_co_timeline)
```


### Harmattan vs non-Harmattan + Cooktime vs non-Cooktime

```{r co trends harmattan and cooktime calc, echo=FALSE, message=FALSE, warning=FALSE}
harmattan_start <- as.Date("2023-12-01")
harmattan_end <- as.Date("2024-03-01")

co_community_hourly_seasonality <- co_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 16 & hour <= 19 ~ "Evening Cooking Hours", #4-9pm
      hour >= 5 & hour <= 8 ~ "Morning Cooking Hours", #5-9am
      (hour < 16 | hour > 19) &  (hour < 5 | hour > 8) & hour != 3 ~ "Other Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Evening Cooking Hours", "Morning Cooking Hours", "Control Hour", "Other Hours"))
  )


# hourly pm 2.5 data from just harmattan only
harmattan_co_hourly <- co_community_hourly_seasonality %>% filter(date >= harmattan_start & date <= harmattan_end) %>% 
  filter(!is.na(hour)) 

# hourly pm 2.5 data from not harmattan only
not_harmattan_co_hourly <- co_community_hourly_seasonality %>% filter(date < harmattan_start | date > harmattan_end) %>%
  filter(!is.na(hour)) 


# mean and sd co during harmattan
mean(harmattan_co_hourly$mean_co, na.rm = TRUE)
sd(harmattan_co_hourly$mean_co, na.rm = TRUE)

# mean and sd co NOT during harmattan
mean(not_harmattan_co_hourly$mean_co, na.rm = TRUE)
sd(not_harmattan_co_hourly$mean_co, na.rm = TRUE)




# COMPARING COOKTIME PM 2.5 ACROSS ALL SEASONS
anova_result <- aov(mean_co ~ cooking_period, data = co_community_hourly_seasonality)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)



# COMPARING COOKTIME PM 2.5 DURING HARMATTAN

anova_result <- aov(mean_co ~ cooking_period, data = harmattan_co_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)



# COMPARING COOKTIME PM 2.5 NOT HARMATTAN

anova_result <- aov(mean_co ~ cooking_period, data = not_harmattan_co_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)




# COOKTIME VS NOT COOKTIME COMP (MEANS AND SDs)

evening_cooktime_co_hourly <- co_community_hourly_seasonality %>% filter(cooking_period == "Evening Cooking Hours") %>% 
  filter(!is.na(hour)) 

morning_cooktime_co_hourly <- co_community_hourly_seasonality %>% filter(cooking_period == "Morning Cooking Hours") %>% 
  filter(!is.na(hour)) 

other_co_hourly <- co_community_hourly_seasonality %>% filter(cooking_period == "Other Hours") %>% 
  filter(!is.na(hour)) 

control_co_hourly <- co_community_hourly_seasonality %>% filter(cooking_period == "Control Hour") %>% 
  filter(!is.na(hour)) 

mean(evening_cooktime_co_hourly$mean_co, na.rm = TRUE) 
sd(evening_cooktime_co_hourly$mean_co, na.rm = TRUE) 

mean(morning_cooktime_co_hourly$mean_co, na.rm = TRUE)  
sd(morning_cooktime_co_hourly$mean_co, na.rm = TRUE) 

mean(other_co_hourly$mean_co, na.rm = TRUE)  
sd(other_co_hourly$mean_co, na.rm = TRUE) 

mean(control_co_hourly$mean_co, na.rm = TRUE)  
sd(control_co_hourly$mean_co, na.rm = TRUE)


# COOKTIME VS NOT COOKTIME VS CONTROL DURING HARMATTAN (MEANS AND SDs)

mean(harmattan_co_hourly$mean_co[harmattan_co_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)
sd(harmattan_co_hourly$mean_co[harmattan_co_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)

mean(harmattan_co_hourly$mean_co[harmattan_co_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)
sd(harmattan_co_hourly$mean_co[harmattan_co_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)

mean(harmattan_co_hourly$mean_co[harmattan_co_hourly$cooking_period == "Other Hours"], na.rm = TRUE)
sd(harmattan_co_hourly$mean_co[harmattan_co_hourly$cooking_period == "Other Hours"], na.rm = TRUE)

mean(harmattan_co_hourly$mean_co[harmattan_co_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(harmattan_co_hourly$mean_co[harmattan_co_hourly$cooking_period == "Control Hour"], na.rm = TRUE)

# COOKTIME VS NOT COOKTIME NOT DURING HARMATTAN (MEANS AND SDs)

mean(not_harmattan_co_hourly$mean_co[not_harmattan_co_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_co_hourly$mean_co[not_harmattan_co_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_co_hourly$mean_co[not_harmattan_co_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_co_hourly$mean_co[not_harmattan_co_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_co_hourly$mean_co[not_harmattan_co_hourly$cooking_period == "Other Hours"], na.rm = TRUE)
sd(not_harmattan_co_hourly$mean_co[not_harmattan_co_hourly$cooking_period == "Other Hours"], na.rm = TRUE)

mean(not_harmattan_co_hourly$mean_co[not_harmattan_co_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(not_harmattan_co_hourly$mean_co[not_harmattan_co_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
```

```{r co trends harmattan and cooktime plot, echo=FALSE, message=FALSE, warning=FALSE}
## PLOTTING THE COMPARISONS 

base_plot <- co_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 16 & hour <= 19 ~ "Evening Cooking Hours", #4-9pm
      hour >= 5 & hour <= 8 ~ "Morning Cooking Hours", #5-9am
      (hour < 16 | hour > 19) &  (hour < 5 | hour > 8) & hour != 3 ~ "Other Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Evening Cooking Hours", "Morning Cooking Hours", "Control Hour", "Other Hours"))
  ) %>%
  filter(!is.na(hour)) %>%
  ggplot(aes(x = harmattan, y = mean_co, fill = cooking_period)) +
  geom_boxplot() +
  theme_classic() +
  labs(y = "CO Concentration (ppb)", 
       x = "Season",
       fill = "Cooking\nPeriod") +  # Change legend title and add line breaks
  theme(legend.position = "right") +
  scale_fill_manual(values = c(
    "Evening Cooking Hours" = "goldenrod", 
    "Morning Cooking Hours" = "#7db569", 
    "Control Hour" = "#b672e0",
    "Other Hours" = "#52b3eb"
  ),
  labels = c("Evening Cooking Hours (4pm-8pm)", "Morning Cooking Hours (5am-9am)", "Control Hour (3am)", "Other Hours (Excluding Cooking \nand Control Hours)"))  # Add line breaks in legend labels



## NOTE: SIGNFICANCE STARS AND BARS CREATED MANUALLY USING ANOVA AND TUKEY TESTS ABOVE. NEW DATA MAY REQURE DIFFERENT STARS.
base_plot +
  geom_segment(aes(x = 0.72, xend = 0.90, y = 10000, yend = 10000), color = "black") +  # Harmattan 
  annotate("text", x = 0.81, y = 10100, label = "***") +
  
  geom_segment(aes(x = 0.91, xend = 1.09, y = 10000, yend = 10000), color = "black") +  # Harmattan 
  annotate("text", x = 1, y = 10100, label = "***") +
  
  geom_segment(aes(x = 1.10, xend = 1.28, y = 10000, yend = 10000), color = "black") +  # Harmattan 
  annotate("text", x = 1.19, y = 10100, label = "") +
  
  geom_segment(aes(x = 0.72, xend = 1.09, y = 10250, yend = 10250), color = "black") +  # Harmattan 
  annotate("text", x = 0.9, y = 10350, label = "***") +
  
  geom_segment(aes(x = 0.91, xend = 1.27, y = 10550, yend = 10550), color = "black") +  # Harmattan
  annotate("text", x = 1.09, y = 10675, label = "***") +
  
  geom_segment(aes(x = 0.72, xend = 1.27, y = 10850, yend = 10850), color = "black") +  # Harmattan
  annotate("text", x = 1, y = 11000, label = "***") +
  
  
  
  geom_segment(aes(x = 1.72, xend = 1.90, y = 10000, yend = 10000), color = "black") +  # Not Harmattan 
  annotate("text", x = 1.81, y = 10100, label = "***") +
  
  geom_segment(aes(x = 1.91, xend = 2.09, y = 10000, yend = 10000), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.0, y = 10100, label = "***") +
  
  geom_segment(aes(x = 2.10, xend = 2.28, y = 10000, yend = 10000), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.19, y = 10100, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 2.09, y = 10250, yend = 10250), color = "black") +  # Not Harmattan 
  annotate("text", x = 1.9, y = 10350, label = "***") +
  
  geom_segment(aes(x = 1.91, xend = 2.28, y = 10550, yend = 10550), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.09, y = 10675, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 2.27, y = 10850, yend = 10850), color = "black") +  # Not Harmattan 
  annotate("text", x = 2, y = 11000, label = "***") +
  
  
  geom_segment(aes(x = 1, xend = 2, y = 11300, yend = 11300), color = "black") +  # Harmattan vs Not Harmattan
  annotate("text", x = 1.5, y = 11425, label = "***") +
  
  labs(caption ="Comparisons and their significance are denoted by the horizontal lines, where * indicates p <0.05, ** indicates p < 0.01, and *** indicates p < 0.001") +
  
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    legend.text = element_text(size = 10), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    plot.caption = element_text(size = 11),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
  )

```

### Spatial Correlation

```{r spatial correlation 25, echo=FALSE, message=FALSE, warning=FALSE}
generate_spatial_pollution_map(gas_corrected_full, "co", missing_monitor_location)
```

### Making maps for each size class

```{r pm maps during harmattan, echo=FALSE, message=FALSE, warning=FALSE}
# DURING HARMATTAN ----

# CO
co_summary <- gas_corrected_full %>%
  filter(date >= harmattan_start & date < harmattan_end) %>%
  group_by(monitor) %>%
  summarise(mean_co = mean(co, na.rm = TRUE),
            median_co = median(co, na.rm = TRUE)) %>%
  right_join(monitor_points)


ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5)  +
  
  # Add road data
  geom_sf(data = roads_filtered, color = "gray", size = 0.3) +  # Add roads data
  
  # Add monitor points with transparency within the bounding box
  geom_sf(data = co_summary %>% filter(!is.na(mean_co)), aes(geometry = geometry, color = mean_co), alpha = 0.7, size = 6) +

  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  
  # Customize
  labs(title = "Mean CO for Each Monitor During Harmattan",
       x = "", 
       y = "") +
  theme_bw() +
  scale_color_gradient(low = "#0540a1", high = "#eb4c2d",
                       breaks = c(min(co_summary$mean_co), 
                                  min(co_summary$mean_co + (max(co_summary$mean_co) - min(co_summary$mean_co))/2),
                                  max(co_summary$mean_co)),
                       labels = c(paste(round(min(co_summary$mean_co),0)), 
                                  paste(round((min(co_summary$mean_co + (max(co_summary$mean_co) - min(co_summary$mean_co))/2)),0)),
                                  paste(round(max(co_summary$mean_co),0))),
                       name = "CO \n(ppb)") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    plot.title = element_text(size = 14), # Increase plot title size
    legend.text = element_text(size = 12), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    legend.key.size = unit(1.3, "lines") # Increase legend tick mark size
  )
`


```

```{r pm maps NOT during harmattan, echo=FALSE, message=FALSE, warning=FALSE}
# NOT DURING HARMATTTAN ----

co_summary <- gas_corrected_full %>%
  filter(date < harmattan_start | date >= harmattan_end) %>%
  group_by(monitor) %>%
  summarise(mean_co = mean(co, na.rm = TRUE),
            median_co = median(co, na.rm = TRUE)) %>%
  right_join(monitor_points)


ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5)  +
  
  # Add road data
  geom_sf(data = roads_filtered, color = "gray", size = 0.3) +  # Add roads data
  
  # Add monitor points with transparency within the bounding box
  geom_sf(data = co_summary %>% filter(!is.na(mean_co)), aes(geometry = geometry, color = mean_co), alpha = 0.7, size = 6) +

  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  
  # Customize
  labs(title = "Mean CO for Each Monitor Outside of Harmattan",
       x = "", 
       y = "") +
  theme_bw() +
  scale_color_gradient(low = "#0540a1", high = "#eb4c2d",
                       breaks = c(min(co_summary$mean_co), 
                                  min(co_summary$mean_co + (max(co_summary$mean_co) - min(co_summary$mean_co))/2),
                                  max(co_summary$mean_co)),
                       labels = c(paste(round(min(co_summary$mean_co),0)), 
                                  paste(round((min(co_summary$mean_co + (max(co_summary$mean_co) - min(co_summary$mean_co))/2)),0)),
                                  paste(round(max(co_summary$mean_co),0))),
                       name = "CO \n(ppb)") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    plot.title = element_text(size = 14), # Increase plot title size
    legend.text = element_text(size = 12), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    legend.key.size = unit(1.3, "lines") # Increase legend tick mark size
  )
```


## HEATMAP CUSTOMIZATION

```{r heatmap customization, echo = FALSE, warning=FALSE, message=FALSE}
## CO
 
co_community_hourly <- co_community_hourly %>%
    mutate(date = ymd(date))  # Use lubridate to ensure date conversion
  
heat_map_data <- co_community_hourly %>%
    group_by(date, hour) %>%
    summarize(mean_co = mean(mean_co, na.rm = TRUE), .groups = 'drop') %>%
    ungroup() %>%
    mutate(day = day(date),
           month = factor(format(date, "%b"), levels = c("Sep", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug")),
           year = year(date)) %>%
  mutate(log_mean_co = log(mean_co))


ggplot(heat_map_data, aes(day, hour, fill = mean_co)) +
    geom_tile(color = NA, size = 0) + 
    scale_fill_gradientn(
        colors = c("#221f26", "#3f2363", "#872CA2", "#D44292", "#F98477", "#EDD9A1", "#fffceb"), # Specify your custom colors
        values = scales::rescale(c(0, 5, 12, 20, 40, 60, 100)), # Set breaks; adjust these values based on your data range
        breaks = c(200, 2000, 4000, 6000, 7900),  # Add more breaks
        labels = c("200", "2000", "4000", "6000", "7900"),  # Corresponding labels
        name = "CO \n(ppb)"
    ) + 
    facet_grid(~month) +
    scale_y_continuous(trans = "reverse", breaks = unique(heat_map_data$hour)) +
    scale_x_continuous(breaks = c(1, 10, 20, 31)) +
    theme_minimal(base_size = 8) +
    labs(x = "Day",
         y = "Hour") + 
    theme(legend.position = "right",
          plot.title = element_text(size = 14, hjust = 0),
          axis.text.y = element_text(size = 10),
          strip.background = element_rect(colour = "white"),
          strip.text = element_text(size = 13),   # Increase facet group text size
          axis.ticks = element_blank(),
          axis.text = element_text(size = 10),
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 11),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14),
          panel.background = element_blank())
```


## APPENDIX PLOTS

```{r time series co individual monitor locations facet, echo=FALSE, warning= FALSE, message=FALSE}
monitor_names <- tibble::tribble(
  ~monitor, ~description, ~location,
  "MOD-00399", "Dwenewoho", "Dwenewoho GH",
  "MOD-00398", "Kurawura Akura", "Kurawura Akura GH",
  "MOD-00397", "Apesika", "Apesika GH",
  "MOD-00401", "New Longoro", "New Longoro GH",
  "MOD-00400", "Kintampo-PTC Area (Magazine)", "Kintampo GH"
)

monitor_names <- monitor_names %>%
  mutate(location = gsub(" GH", "", location)) %>%
  mutate(locations = if_else(grepl("Kintampo", location),
                            paste(description, location, sep = ", "),
                            location))

# Custom function to add line breaks at a specific length
add_line_breaks <- function(x, width) {
  sapply(x, function(y) {
    if (nchar(y) > width) {
      paste(strwrap(y, width = width), collapse = "\n")
    } else {
      y
    }
  })
}

# Apply custom function to locations
monitor_names <- monitor_names %>%
  mutate(locations = if_else(grepl("Kintampo", location),
                             paste(description, "Kintampo", sep = "\n"),
                             location)) %>%
  mutate(locations = add_line_breaks(locations, width = 20))

# Plot with updated facet labels
gas_corrected_full %>%
  left_join(monitor_names) %>%
  group_by(date, locations) %>%
  summarize(mean_co = mean(co, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_co)) +
  geom_line(color = "black", alpha = 1) +
  theme_minimal() +
  facet_wrap(~locations) +
  labs(y = "Mean CO") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 11)  # Increase facet group text size
  )
```


# Ozone (O3) Community Deployment

### Checking Individual Monitor to Fleet Average

```{r monitor vs fleet avg 25, fig.height=10, fig.width=7, echo=FALSE, message=FALSE, warning=FALSE}
# hourly comparison
compare_fleet_avg_monitor(o3_community_hourly, "o3", "hourly")

# Daily comparison
compare_fleet_avg_monitor(o3_community_daily, "o3", "daily")
```

### Ozone Trends

```{r ozone timelines, echo=FALSE, message=FALSE, warning=FALSE}
daily_o3_all_devices <- gas_corrected_full %>%
  group_by(date, monitor) %>%
  summarize(mean_o3 = mean(o3, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_o3, group = monitor)) +
  geom_line(color = "black", alpha = 0.2) +
  theme_minimal() +
  labs(y = "Mean O3",
       title = "Average O3 Accross All Devices")

daily_cooktime_o3_all_devices <- gas_corrected_full %>%
  filter(hour >= 16 & hour < 20) %>%
  group_by(monitor, date) %>%
  summarise(mean_o3 = mean(o3, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_o3, group = monitor)) +
  geom_line(color = "black", alpha = 0.4) +
  theme_minimal() +
  labs(y = "Mean O3",
       title = "Daily Average O3 for Each Monitor During Primary Cooking Hours (4-8)") 

grid.arrange(daily_o3_all_devices, daily_cooktime_o3_all_devices)

daily_o3_timeline <- gas_corrected_full %>%
  group_by(date) %>%
  summarize(mean_o3 = mean(o3, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_o3)) +
  geom_line() +
  theme_minimal() +
  labs(y = "Mean PM O3",
       title = "Average O3 Accross All Devices")

daily_cooktime_o3_timeline <- gas_corrected_full %>%
  filter(hour >= 16 & hour < 20) %>%
  group_by(date) %>%
  summarize(mean_o3 = mean(o3, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_o3)) +
  geom_line() +
  theme_minimal() +
  labs(y = "Mean O3",
       title = "Mean O3 Accross All Devices During Cooking Hours (4-8)")

grid.arrange(daily_o3_timeline, daily_cooktime_o3_timeline)
```


### Harmattan vs non-Harmattan + Cooktime vs non-Cooktime

```{r ozone trends harmattan and cooktime calc, echo=FALSE, message=FALSE, warning=FALSE}
harmattan_start <- as.Date("2023-12-01")
harmattan_end <- as.Date("2024-03-01")

o3_community_hourly_seasonality <- o3_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 15 & hour <= 18 ~ "Active Cooking Hours",
      (hour < 15 | hour > 18) & hour != 3 ~ "Non-Active Cooking Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Active Cooking Hours", "Non-Active Cooking Hours", "Control Hour"))
  )


# hourly pm 2.5 data from just harmattan only
harmattan_o3_hourly <- o3_community_hourly_seasonality %>% filter(date >= harmattan_start & date <= harmattan_end) %>% 
  filter(!is.na(hour)) 

# hourly pm 2.5 data from not harmattan only
not_harmattan_o3_hourly <- o3_community_hourly_seasonality %>% filter(date < harmattan_start | date > harmattan_end) %>%
  filter(!is.na(hour)) 


# mean and sd o3 during harmattan
mean(harmattan_o3_hourly$mean_o3, na.rm = TRUE)
sd(harmattan_o3_hourly$mean_o3, na.rm = TRUE)

# mean and sd o3 NOT during harmattan
mean(not_harmattan_o3_hourly$mean_o3, na.rm = TRUE)
sd(not_harmattan_o3_hourly$mean_o3, na.rm = TRUE)




# COMPARING COOKTIME O3 ACROSS ALL SEASONS
anova_result <- aov(mean_o3 ~ cooking_period, data = o3_community_hourly_seasonality)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)



# COMPARING COOKTIME PM 2.5 DURING HARMATTAN

anova_result <- aov(mean_o3 ~ cooking_period, data = harmattan_o3_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)



# COMPARING COOKTIME O3 NOT HARMATTAN

anova_result <- aov(mean_o3 ~ cooking_period, data = not_harmattan_o3_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)




# COOKTIME VS NOT COOKTIME COMP (MEANS AND SDs)

cooktime_o3_hourly <- o3_community_hourly_seasonality %>% filter(cooking_period == "Active Cooking Hours") %>% 
  filter(!is.na(hour)) 

not_cooktime_o3_hourly <- o3_community_hourly_seasonality %>% filter(cooking_period == "Non-Active Cooking Hours") %>% 
  filter(!is.na(hour)) 

control_o3_hourly <- o3_community_hourly_seasonality %>% filter(cooking_period == "Control Hour") %>% 
  filter(!is.na(hour)) 

mean(cooktime_o3_hourly$mean_o3, na.rm = TRUE) 
sd(cooktime_o3_hourly$mean_o3, na.rm = TRUE) 

mean(not_cooktime_o3_hourly$mean_o3, na.rm = TRUE)  
sd(not_cooktime_o3_hourly$mean_o3, na.rm = TRUE) 

mean(control_o3_hourly$mean_o3, na.rm = TRUE)  
sd(control_o3_hourly$mean_o3, na.rm = TRUE)


# COOKTIME VS NOT COOKTIME VS CONTROL DURING HARMATTAN (MEANS AND SDs)

mean(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Active Cooking Hours"], na.rm = TRUE)
sd(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Active Cooking Hours"], na.rm = TRUE)

mean(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Non-Active Cooking Hours"], na.rm = TRUE)
sd(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Non-Active Cooking Hours"], na.rm = TRUE)

mean(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Control Hour"], na.rm = TRUE)

# COOKTIME VS NOT COOKTIME NOT DURING HARMATTAN (MEANS AND SDs)

mean(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Active Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Active Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Non-Active Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Non-Active Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
```

```{r co trends harmattan and cooktime plot, echo=FALSE, message=FALSE, warning=FALSE}
## PLOTTING THE COMPARISONS 

base_plot <- o3_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 15 & hour <= 18 ~ "Active Cooking Hours",
      (hour < 15 | hour > 18) & hour != 3 ~ "Non-Active Cooking Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Active Cooking Hours", "Non-Active Cooking Hours", "Control Hour"))
  ) %>%
  filter(!is.na(hour)) %>%
  ggplot(aes(x = harmattan, y = mean_o3, fill = cooking_period)) +
  geom_boxplot() +
  theme_classic() +
  labs(y = "O3 Concentration (ppb)", 
       x = "Season",
       fill = "Cooking\nPeriod") +  # Change legend title and add line breaks
  theme(legend.position = "right") +
  scale_fill_manual(values = c(
    "Active Cooking Hours" = "goldenrod", 
    "Non-Active Cooking Hours" = "#52b3eb", 
    "Control Hour" = "#b672e0"
  ),
  labels = c("Evening Cooking Hours (4pm-8pm)", "Other Hours (Excluding Evening Cooking \nand Control Hours)", "Control Hour (3am)"))  # Add line breaks in legend labels

## NOTE: SIGNFICANCE STARS AND BARS CREATED MANUALLY USING ANOVA AND TUKEY TESTS ABOVE. NEW DATA MAY REQURE DIFFERENT STARS.
base_plot +
  geom_segment(aes(x = 0.72, xend = 0.98, y = 122, yend = 122), color = "black") +  # Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 0.85, y = 125, label = "***") +
  
  geom_segment(aes(x = 1.02, xend = 1.28, y = 122, yend = 122), color = "black") +  # Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 1.15, y = 125, label = "***") +
  
  geom_segment(aes(x = 0.72, xend = 1.27, y = 128, yend = 128), color = "black") +  # Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 1, y = 130, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 1.98, y = 122, yend = 122), color = "black") +  # Not Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 1.85, y = 125, label = "***") +
  
  geom_segment(aes(x = 2.02, xend = 2.28, y = 122, yend = 122), color = "black") +  # Not Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 2.15, y = 125, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 2.27, y = 128, yend = 128), color = "black") +  # Not Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 2, y = 130, label = "***") +
  
  geom_segment(aes(x = 1, xend = 2, y = 133, yend = 133), color = "black") +  # Harmattan vs Not Harmattan
  annotate("text", x = 1.5, y = 135, label = "***") +
  
  labs(caption ="Comparisons and their significance are denoted by the horizontal lines, where the three stars indicate p < 0.001.") +
  
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    legend.text = element_text(size = 10), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    plot.caption = element_text(size = 11),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
  )
```


#### Adding morning cooking hours

```{r ozone trends harmattan and cooktime calc, echo=FALSE, message=FALSE, warning=FALSE}
harmattan_start <- as.Date("2023-12-01")
harmattan_end <- as.Date("2024-03-01")

o3_community_hourly_seasonality <- o3_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 16 & hour <= 19 ~ "Evening Cooking Hours", #4-9pm
      hour >= 5 & hour <= 8 ~ "Morning Cooking Hours", #5-9am
      (hour < 16 | hour > 19) &  (hour < 5 | hour > 8) & hour != 3 ~ "Other Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Evening Cooking Hours", "Morning Cooking Hours", "Control Hour", "Other Hours"))
  )


# hourly pm 2.5 data from just harmattan only
harmattan_o3_hourly <- o3_community_hourly_seasonality %>% filter(date >= harmattan_start & date <= harmattan_end) %>% 
  filter(!is.na(hour)) 

# hourly pm 2.5 data from not harmattan only
not_harmattan_o3_hourly <- o3_community_hourly_seasonality %>% filter(date < harmattan_start | date > harmattan_end) %>%
  filter(!is.na(hour)) 


# mean and sd o3 during harmattan
mean(harmattan_o3_hourly$mean_o3, na.rm = TRUE)
sd(harmattan_o3_hourly$mean_o3, na.rm = TRUE)

# mean and sd o3 NOT during harmattan
mean(not_harmattan_o3_hourly$mean_o3, na.rm = TRUE)
sd(not_harmattan_o3_hourly$mean_o3, na.rm = TRUE)




# COMPARING COOKTIME O3 ACROSS ALL SEASONS
anova_result <- aov(mean_o3 ~ cooking_period, data = o3_community_hourly_seasonality)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)



# COMPARING COOKTIME O3 DURING HARMATTAN

anova_result <- aov(mean_o3 ~ cooking_period, data = harmattan_o3_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)



# COMPARING COOKTIME O3 NOT HARMATTAN

anova_result <- aov(mean_o3 ~ cooking_period, data = not_harmattan_o3_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)




# COOKTIME VS NOT COOKTIME COMP (MEANS AND SDs)

evening_cooktime_o3_hourly <- o3_community_hourly_seasonality %>% filter(cooking_period == "Evening Cooking Hours") %>% 
  filter(!is.na(hour)) 

morning_cooktime_o3_hourly <- o3_community_hourly_seasonality %>% filter(cooking_period == "Morning Cooking Hours") %>% 
  filter(!is.na(hour)) 

other_o3_hourly <- o3_community_hourly_seasonality %>% filter(cooking_period == "Other Hours") %>% 
  filter(!is.na(hour)) 

control_o3_hourly <- o3_community_hourly_seasonality %>% filter(cooking_period == "Control Hour") %>% 
  filter(!is.na(hour)) 


mean(evening_cooktime_o3_hourly$mean_o3, na.rm = TRUE) 
sd(evening_cooktime_o3_hourly$mean_o3, na.rm = TRUE) 

mean(morning_cooktime_o3_hourly$mean_o3, na.rm = TRUE)  
sd(morning_cooktime_o3_hourly$mean_o3, na.rm = TRUE) 

mean(other_o3_hourly$mean_o3, na.rm = TRUE)  
sd(other_o3_hourly$mean_o3, na.rm = TRUE) 

mean(control_o3_hourly$mean_o3, na.rm = TRUE)  
sd(control_o3_hourly$mean_o3, na.rm = TRUE)


# COOKTIME VS NOT COOKTIME VS CONTROL DURING HARMATTAN (MEANS AND SDs)

mean(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)
sd(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)

mean(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)
sd(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)

mean(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Other Hours"], na.rm = TRUE)
sd(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Other Hours"], na.rm = TRUE)

mean(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(harmattan_o3_hourly$mean_o3[harmattan_o3_hourly$cooking_period == "Control Hour"], na.rm = TRUE)

# COOKTIME VS NOT COOKTIME NOT DURING HARMATTAN (MEANS AND SDs)

mean(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Other Hours"], na.rm = TRUE)
sd(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Other Hours"], na.rm = TRUE)

mean(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(not_harmattan_o3_hourly$mean_o3[not_harmattan_o3_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
```

```{r co trends harmattan and cooktime plot, echo=FALSE, message=FALSE, warning=FALSE}
## PLOTTING THE COMPARISONS 

base_plot <- o3_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 16 & hour <= 19 ~ "Evening Cooking Hours", #4-9pm
      hour >= 5 & hour <= 8 ~ "Morning Cooking Hours", #5-9am
      (hour < 16 | hour > 19) &  (hour < 5 | hour > 8) & hour != 3 ~ "Other Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Evening Cooking Hours", "Morning Cooking Hours", "Control Hour", "Other Hours"))
  ) %>%
  filter(!is.na(hour)) %>%
  ggplot(aes(x = harmattan, y = mean_o3, fill = cooking_period)) +
  geom_boxplot() +
  theme_classic() +
  labs(y = "O3 Concentration (ppb)", 
       x = "Season",
       fill = "Cooking\nPeriod") +  # Change legend title and add line breaks
  theme(legend.position = "right") +
  scale_fill_manual(values = c(
    "Evening Cooking Hours" = "goldenrod", 
    "Morning Cooking Hours" = "#7db569", 
    "Control Hour" = "#b672e0",
    "Other Hours" = "#52b3eb"
  ),
  labels = c("Evening Cooking Hours (4pm-8pm)", "Morning Cooking Hours (5am-9am)", "Control Hour (3am)", "Other Hours (Excluding Cooking \nand Control Hours)"))  # Add line breaks in legend labels


## NOTE: SIGNFICANCE STARS AND BARS CREATED MANUALLY USING ANOVA AND TUKEY TESTS ABOVE. NEW DATA MAY REQURE DIFFERENT STARS.
base_plot +
  geom_segment(aes(x = 0.72, xend = 0.90, y = 121, yend = 121), color = "black") +  # Harmattan 
  annotate("text", x = 0.81, y = 122, label = "***") +
  
  geom_segment(aes(x = 0.91, xend = 1.09, y = 121, yend = 121), color = "black") +  # Harmattan 
  annotate("text", x = 1, y = 122, label = "***") +
  
  geom_segment(aes(x = 1.10, xend = 1.28, y = 121, yend = 121), color = "black") +  # Harmattan 
  annotate("text", x = 1.19, y = 122, label = "**") +
  
  geom_segment(aes(x = 0.72, xend = 1.09, y = 125, yend = 125), color = "black") +  # Harmattan 
  annotate("text", x = 0.9, y = 126, label = "***") +
  
  geom_segment(aes(x = 0.91, xend = 1.27, y = 129, yend = 129), color = "black") +  # Harmattan
  annotate("text", x = 1.09, y = 130, label = "***") +
  
  geom_segment(aes(x = 0.72, xend = 1.27, y = 133, yend = 133), color = "black") +  # Harmattan
  annotate("text", x = 1, y = 135, label = "***") +
  
  
  
  geom_segment(aes(x = 1.72, xend = 1.90, y = 121, yend = 121), color = "black") +  # Not Harmattan 
  annotate("text", x = 1.81, y = 122, label = "***") +
  
  geom_segment(aes(x = 1.91, xend = 2.09, y = 121, yend = 121), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.0, y = 122, label = "") +
  
  geom_segment(aes(x = 2.10, xend = 2.28, y = 121, yend = 121), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.19, y = 122, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 2.09, y = 125, yend = 125), color = "black") +  # Not Harmattan 
  annotate("text", x = 1.9, y = 126, label = "***") +
  
  geom_segment(aes(x = 1.91, xend = 2.28, y = 129, yend = 129), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.09, y = 130, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 2.27, y = 133, yend = 133), color = "black") +  # Not Harmattan 
  annotate("text", x = 2, y = 135, label = "***") +
  
  
  geom_segment(aes(x = 1, xend = 2, y = 140, yend = 140), color = "black") +  # Harmattan vs Not Harmattan
  annotate("text", x = 1.5, y = 142, label = "***") +
  
  labs(caption ="Comparisons and their significance are denoted by the horizontal lines, where * indicates p <0.05, ** indicates p < 0.01, and *** indicates p < 0.001.") +
  
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    legend.text = element_text(size = 10), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    plot.caption = element_text(size = 11),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
  )
```



### Spatial Correlation

```{r spatial correlation 25, echo=FALSE, message=FALSE, warning=FALSE}
generate_spatial_pollution_map(gas_corrected_full, "o3", missing_monitor_location)
```

### Making maps for each size class

```{r pm maps during harmattan, echo=FALSE, message=FALSE, warning=FALSE}
# DURING HARMATTAN ----

# O3
o3_summary <- gas_corrected_full %>%
  filter(date >= harmattan_start & date < harmattan_end) %>%
  group_by(monitor) %>%
  summarise(mean_o3 = mean(o3, na.rm = TRUE),
            median_o3 = median(o3, na.rm = TRUE)) %>%
  right_join(monitor_points)


ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5)  +
  
  # Add road data
  geom_sf(data = roads_filtered, color = "gray", size = 0.3) +  # Add roads data
  
  # Add monitor points with transparency within the bounding box
  geom_sf(data = o3_summary %>% filter(!is.na(mean_o3)), aes(geometry = geometry, color = mean_o3), alpha = 0.7, size = 6) +

  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  
  # Customize
  labs(title = "Mean O3 for Each Monitor During Harmattan",
       x = "", 
       y = "") +
  theme_bw() +
  scale_color_gradient(low = "#0540a1", high = "#eb4c2d",
                       breaks = c(min(o3_summary$mean_o3), 
                                  min(o3_summary$mean_o3 + (max(o3_summary$mean_o3) - min(o3_summary$mean_o3))/2),
                                  max(o3_summary$mean_o3)),
                       labels = c(paste(round(min(o3_summary$mean_o3),0)), 
                                  paste(round((min(o3_summary$mean_o3 + (max(o3_summary$mean_o3) - min(o3_summary$mean_o3))/2)),0)),
                                  paste(round(max(o3_summary$mean_o3),0))),
                       name = "O3 \n(ppb)") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    plot.title = element_text(size = 14), # Increase plot title size
    legend.text = element_text(size = 12), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    legend.key.size = unit(1.3, "lines") # Increase legend tick mark size
  )
`


```

```{r pm maps NOT during harmattan, echo=FALSE, message=FALSE, warning=FALSE}
# NOT DURING HARMATTTAN ----

o3_summary <- gas_corrected_full %>%
  filter(date < harmattan_start | date >= harmattan_end) %>%
  group_by(monitor) %>%
  summarise(mean_o3 = mean(o3, na.rm = TRUE),
            median_o3 = median(o3, na.rm = TRUE)) %>%
  right_join(monitor_points)


ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5)  +
  
  # Add road data
  geom_sf(data = roads_filtered, color = "gray", size = 0.3) +  # Add roads data
  
  # Add monitor points with transparency within the bounding box
  geom_sf(data = o3_summary %>% filter(!is.na(mean_o3)), aes(geometry = geometry, color = mean_o3), alpha = 0.7, size = 6) +

  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  
  # Customize
  labs(title = "Mean O3 for Each Monitor Outside of Harmattan",
       x = "", 
       y = "") +
  theme_bw() +
  scale_color_gradient(low = "#0540a1", high = "#eb4c2d",
                       breaks = c(min(o3_summary$mean_o3), 
                                  min(o3_summary$mean_o3 + (max(o3_summary$mean_o3) - min(o3_summary$mean_o3))/2),
                                  max(o3_summary$mean_o3)),
                       labels = c(paste(round(min(o3_summary$mean_o3),0)), 
                                  paste(round((min(o3_summary$mean_o3 + (max(o3_summary$mean_o3) - min(o3_summary$mean_o3))/2)),0)),
                                  paste(round(max(o3_summary$mean_o3),0))),
                       name = "O3 \n(ppb)") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    plot.title = element_text(size = 14), # Increase plot title size
    legend.text = element_text(size = 12), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    legend.key.size = unit(1.3, "lines") # Increase legend tick mark size
  )
```


## HEATMAP CUSTOMIZATION

```{r heatmap customization, echo = FALSE, warning=FALSE, message=FALSE}
## O3
 
o3_community_hourly <- o3_community_hourly %>%
    mutate(date = ymd(date))  # Use lubridate to ensure date conversion
  
heat_map_data <- o3_community_hourly %>%
    group_by(date, hour) %>%
    summarize(mean_o3 = mean(mean_o3, na.rm = TRUE), .groups = 'drop') %>%
    ungroup() %>%
    mutate(day = day(date),
           month = factor(format(date, "%b"), levels = c("Sep", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug")),
           year = year(date)) %>%
  mutate(log_mean_o3 = log(mean_o3))


ggplot(heat_map_data, aes(day, hour, fill = mean_o3)) +
    geom_tile(color = NA, size = 0) + 
    scale_fill_gradientn(
        colors = c("#221f26", "#3f2363", "#872CA2", "#D44292", "#F98477", "#EDD9A1", "#fffceb"), # Specify your custom colors
        values = scales::rescale(c(0, 20, 40, 60, 80, 100)), # Set breaks; adjust these values based on your data range
        # breaks = c(25, 50, 75, 100),  # Add more breaks
        # labels = c("25", "50", "75", "100"),  # Corresponding labels
        name = "O3 \n(ppb)"
    ) + 
    facet_grid(~month) +
    scale_y_continuous(trans = "reverse", breaks = unique(heat_map_data$hour)) +
    scale_x_continuous(breaks = c(1, 10, 20, 31)) +
    theme_minimal(base_size = 8) +
    labs(x = "Day",
         y = "Hour") + 
    theme(legend.position = "right",
          plot.title = element_text(size = 14, hjust = 0),
          axis.text.y = element_text(size = 10),
          strip.background = element_rect(colour = "white"),
          strip.text = element_text(size = 13),   # Increase facet group text size
          axis.ticks = element_blank(),
          axis.text = element_text(size = 10),
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 11),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14),
          panel.background = element_blank())
```


## APPENDIX PLOTS

```{r time series ozone individual monitor locations facet, echo=FALSE, warning= FALSE, message=FALSE}
monitor_names <- tibble::tribble(
  ~monitor, ~description, ~location,
  "MOD-00399", "Dwenewoho", "Dwenewoho GH",
  "MOD-00398", "Kurawura Akura", "Kurawura Akura GH",
  "MOD-00397", "Apesika", "Apesika GH",
  "MOD-00401", "New Longoro", "New Longoro GH",
  "MOD-00400", "Kintampo-PTC Area (Magazine)", "Kintampo GH"
)

monitor_names <- monitor_names %>%
  mutate(location = gsub(" GH", "", location)) %>%
  mutate(locations = if_else(grepl("Kintampo", location),
                            paste(description, location, sep = ", "),
                            location))

# Custom function to add line breaks at a specific length
add_line_breaks <- function(x, width) {
  sapply(x, function(y) {
    if (nchar(y) > width) {
      paste(strwrap(y, width = width), collapse = "\n")
    } else {
      y
    }
  })
}

# Apply custom function to locations
monitor_names <- monitor_names %>%
  mutate(locations = if_else(grepl("Kintampo", location),
                             paste(description, "Kintampo", sep = "\n"),
                             location)) %>%
  mutate(locations = add_line_breaks(locations, width = 20))

# Plot with updated facet labels
gas_corrected_full %>%
  left_join(monitor_names) %>%
  group_by(date, locations) %>%
  summarize(mean_o3 = mean(o3, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_o3)) +
  geom_line(color = "black", alpha = 1) +
  theme_minimal() +
  facet_wrap(~locations) +
  labs(y = "Mean O3") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 11)  # Increase facet group text size
  )
```



#### Ozone dual axis plots 

```{r}
## read in temp data 

temp_clean <- read_rds(here("data", "weather", "merged", "met_temp_merged_20230815-20240820.rds")) %>%
  mutate(hour = hour(timestamp)) 

temp_community_hourly <- read_rds(here("data", "weather", "summarized", "temp_hourly_20230815-20240820.rds")) 

temp_community_daily <- read_rds(here("data", "weather", "summarized", "temp_daily_20230815-20240820.rds")) 


## read in PM data

pm25_corrected <- readRDS(here("data", "pm", "final", "pm25corrected_20231024-20240816.rds"))

pm25_community_hourly <- readRDS(here("data", "pm", "summarized", "pm25_community_hourly_20231024-20240816.rds"))
pm25_community_daily <- readRDS(here("data", "pm", "summarized", "pm25_community_daily_20231024-20240816.rds"))

```



```{r}
daily_o3_all_monitors <- gas_corrected_full %>%
  group_by(date) %>%
  summarize(mean_o3 = mean(o3, na.rm = TRUE))

daily_temp_all_monitors <- temp_clean %>%
  group_by(date) %>%
  summarize(mean_temp = mean(met_temp, na.rm = TRUE)) %>%
  filter(date < as.Date("2024-08-17")) %>%
  filter(date > as.Date("2023-09-25"))


daily_pm25_all_monitors <- pm25_corrected %>%
  group_by(date) %>%
  summarize(mean_pm25 = mean(pm25, na.rm = TRUE)) 

ozone_temp_pm_full <- left_join(daily_o3_all_monitors, daily_temp_all_monitors) %>%
  left_join(daily_pm25_all_monitors)

```

```{r}
# Base R plot with dual y-axes (Ozone vs. Temperature)
par(mar = c(5, 4, 4, 4) + 0.3)  # Adjust margins for second axis

plot(ozone_temp_pm_full$date, ozone_temp_pm_full$mean_o3, 
     type = "l", col = "black", lwd = 2, 
     xlab = "Date", ylab = "Ozone (ppb)", 
     ylim = c(min(ozone_temp_pm_full$mean_o3, na.rm = TRUE), 
              max(ozone_temp_pm_full$mean_o3, na.rm = TRUE)))

par(new = TRUE)  # Allow a second plot overlaid

plot(ozone_temp_pm_full$date, ozone_temp_pm_full$mean_temp, 
     type = "l", col = "red", lwd = 2, 
     axes = FALSE, xlab = "", ylab = "", 
     ylim = c(min(ozone_temp_pm_full$mean_temp, na.rm = TRUE), 
              max(ozone_temp_pm_full$mean_temp, na.rm = TRUE)))

axis(side = 4, col.axis = "red")  # Make Temperature y-axis text red
mtext("Temperature (C)", side = 4, line = 3, col = "red")  # Label right y-axis

title("Ozone and Temperature Time Series")




# Base R plot with dual y-axes (Ozone vs. PM2.5)
par(mar = c(5, 4, 4, 4) + 0.3)  # Adjust margins for second axis

plot(ozone_temp_pm_full$date, ozone_temp_pm_full$mean_o3, 
     type = "l", col = "black", lwd = 2, 
     xlab = "Date", ylab = "Ozone (ppb)", 
     ylim = c(min(ozone_temp_pm_full$mean_o3, na.rm = TRUE), 
              max(ozone_temp_pm_full$mean_o3, na.rm = TRUE)))

par(new = TRUE)  # Allow a second plot overlaid

plot(ozone_temp_pm_full$date, ozone_temp_pm_full$mean_pm25, 
     type = "l", col = "#2259e3", lwd = 2, 
     axes = FALSE, xlab = "", ylab = "", 
     ylim = c(min(ozone_temp_pm_full$mean_pm25, na.rm = TRUE), 
              max(ozone_temp_pm_full$mean_pm25, na.rm = TRUE)))

axis(side = 4, col.axis = "#2259e3")  # Make PM2.5 y-axis text green
mtext("PM2.5 (g/m)", side = 4, line = 3, col = "#2259e3")  # Label right y-axis

title("Ozone and PM2.5 Time Series")  # Add title

```

## Splitting by hour for each season 

```{r}
## DURING HARMATTAN ----
o3_hourly_harmattan <- o3_community_hourly %>%
  filter(date > harmattan_start,
         date < harmattan_end) %>%
  group_by(hour) %>%
  summarise(mean_o3 = mean(mean_o3, na.rm = TRUE))

pm25_hourly_harmattan <- pm25_community_hourly %>%
  filter(date > harmattan_start,
         date < harmattan_end) %>%
  group_by(hour) %>%
  summarise(mean_pm25 = mean(mean_pm25, na.rm = TRUE))

temp_hourly_harmattan <- temp_community_hourly %>%
  filter(date > harmattan_start,
         date < harmattan_end) %>%
  group_by(hour) %>%
  summarise(mean_temp = mean(mean_met_temp, na.rm = TRUE))
  
hourly_harmattan_measures <- o3_hourly_harmattan %>% left_join(pm25_hourly_harmattan) %>% left_join(temp_hourly_harmattan)

# Base R plot with dual y-axes (Ozone vs. Temperature)
par(mar = c(5, 4, 4, 4) + 0.3)  # Adjust margins for second axis

plot(hourly_harmattan_measures$hour, hourly_harmattan_measures$mean_o3, 
     type = "l", col = "black", lwd = 2, 
     xlab = "Date", ylab = "Ozone (ppb)", 
     ylim = c(min(hourly_harmattan_measures$mean_o3, na.rm = TRUE), 
              max(hourly_harmattan_measures$mean_o3, na.rm = TRUE)))

par(new = TRUE)  # Allow a second plot overlaid

plot(hourly_harmattan_measures$hour, hourly_harmattan_measures$mean_temp, 
     type = "l", col = "red", lwd = 2, 
     axes = FALSE, xlab = "", ylab = "", 
     ylim = c(min(hourly_harmattan_measures$mean_temp, na.rm = TRUE), 
              max(hourly_harmattan_measures$mean_temp, na.rm = TRUE)))

axis(side = 4, col.axis = "red")  # Make Temperature y-axis text red
mtext("Temperature (C)", side = 4, line = 3, col = "red")  # Label right y-axis

title("Hourly Ozone and Temperature During Harmattan")




# Base R plot with dual y-axes (Ozone vs. Temperature)
par(mar = c(5, 4, 4, 4) + 0.3)  # Adjust margins for second axis

plot(hourly_harmattan_measures$hour, hourly_harmattan_measures$mean_o3, 
     type = "l", col = "black", lwd = 2, 
     xlab = "Date", ylab = "Ozone (ppb)", 
     ylim = c(min(hourly_harmattan_measures$mean_o3, na.rm = TRUE), 
              max(hourly_harmattan_measures$mean_o3, na.rm = TRUE)))

par(new = TRUE)  # Allow a second plot overlaid

plot(hourly_harmattan_measures$hour, hourly_harmattan_measures$mean_pm25, 
     type = "l", col = "#2259e3", lwd = 2, 
     axes = FALSE, xlab = "", ylab = "", 
     ylim = c(min(hourly_harmattan_measures$mean_pm25, na.rm = TRUE), 
              max(hourly_harmattan_measures$mean_pm25, na.rm = TRUE)))

axis(side = 4, col.axis = "#2259e3")  # Make Temperature y-axis text red
mtext("PM2.5 (g/m)", side = 4, line = 3, col = "#2259e3")  # Label right y-axis

title("Hourly Ozone and PM 2.5 During Harmattan")








  ## OUTSIDE OF HARMATTAN ----
o3_hourly_not_harmattan <- o3_community_hourly %>%
  filter(date < harmattan_start | date > harmattan_end) %>%
  group_by(hour) %>%
  summarise(mean_o3 = mean(mean_o3, na.rm = TRUE))

pm25_hourly_not_harmattan<- pm25_community_hourly %>%
  filter(date < harmattan_start | date > harmattan_end) %>%
  group_by(hour) %>%
  summarise(mean_pm25 = mean(mean_pm25, na.rm = TRUE))

temp_hourly_not_harmattan <- temp_community_hourly %>%
  filter(date < harmattan_start | date > harmattan_end) %>%
  group_by(hour) %>%
  summarise(mean_temp = mean(mean_met_temp, na.rm = TRUE))

hourly_not_harmattan_measures <- o3_hourly_not_harmattan %>% left_join(pm25_hourly_not_harmattan) %>% left_join(temp_hourly_not_harmattan)




# Base R plot with dual y-axes (Ozone vs. Temperature)
par(mar = c(5, 4, 4, 4) + 0.3)  # Adjust margins for second axis

plot(hourly_not_harmattan_measures$hour, hourly_not_harmattan_measures$mean_o3, 
     type = "l", col = "black", lwd = 2, 
     xlab = "Date", ylab = "Ozone (ppb)", 
     ylim = c(min(hourly_not_harmattan_measures$mean_o3, na.rm = TRUE), 
              max(hourly_not_harmattan_measures$mean_o3, na.rm = TRUE)))

par(new = TRUE)  # Allow a second plot overlaid

plot(hourly_not_harmattan_measures$hour, hourly_not_harmattan_measures$mean_temp, 
     type = "l", col = "red", lwd = 2, 
     axes = FALSE, xlab = "", ylab = "", 
     ylim = c(min(hourly_not_harmattan_measures$mean_temp, na.rm = TRUE), 
              max(hourly_not_harmattan_measures$mean_temp, na.rm = TRUE)))

axis(side = 4, col.axis = "red")  # Make Temperature y-axis text red
mtext("Temperature (C)", side = 4, line = 3, col = "red")  # Label right y-axis

title("Hourly Ozone and Temperature Outside of Harmattan")


# Base R plot with dual y-axes (Ozone vs. Temperature)
par(mar = c(5, 4, 4, 4) + 0.3)  # Adjust margins for second axis

plot(hourly_not_harmattan_measures$hour, hourly_not_harmattan_measures$mean_o3, 
     type = "l", col = "black", lwd = 2, 
     xlab = "Date", ylab = "Ozone (ppb)", 
     ylim = c(min(hourly_not_harmattan_measures$mean_o3, na.rm = TRUE), 
              max(hourly_not_harmattan_measures$mean_o3, na.rm = TRUE)))

par(new = TRUE)  # Allow a second plot overlaid

plot(hourly_not_harmattan_measures$hour, hourly_not_harmattan_measures$mean_pm25, 
     type = "l", col = "#2259e3", lwd = 2, 
     axes = FALSE, xlab = "", ylab = "", 
     ylim = c(min(hourly_not_harmattan_measures$mean_pm25, na.rm = TRUE), 
              max(hourly_not_harmattan_measures$mean_pm25, na.rm = TRUE)))

axis(side = 4, col.axis = "#2259e3")  # Make Temperature y-axis text red
mtext("PM2.5 (g/m)", side = 4, line = 3, col = "#2259e3")  # Label right y-axis

title("Hourly Ozone and PM 2.5 Outside of Harmattan")

```

