---
title: "Ghana Air Quality Report"
output: 
  html_document:
    toc: true
    theme: united
date: "2024-06-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
### load packages 
library(here) # file path org
library(lubridate)# working with dates
library(tictoc) # timing
library(DT) # datatables
library(purrr) # applying functions across df
library(tidyverse) # data cleaning and plotting
library(data.table) 
library(sf) # spatial data 
library(viridis) # color pallete 
library(knitr)
library(modelsummary) # table of regressions
library(spdep)
library(gstat)
library(units) 
library(gridExtra)
library(broom)
library(Metrics) 
library(kableExtra) # table creation

```


```{r source functions, echo=FALSE, message=FALSE, warning=FALSE}
# Source in functions

# Source function that creates fleet average vs monitor pollutant reading plot
source(here("src", "compare_fleet_avg_monitor.R"))
source(here("src", "compare_fleet_regression.R"))

# Source function that creates heatmap of pollutant readings
source(here("src", "generate_heatmap.R"))

# Source function to make map of pollution values at each monitor location
source(here("src", "generate_spatial_pollution_map.R"))

# source function to calculate moran i 
source(here("src", "calculate_moran_i.R"))
```


```{r location data, echo=FALSE, message=FALSE, warning=FALSE}
#LOAD LOCATION DATA ----
monitor_community_info <- read_csv(here("data", "monitor_community_info.csv"))

# create spatial feature dataset of monitor points
#monitor_points <- st_as_sf(monitor_community_info, coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))


monitor_points <- monitor_community_info %>%
  st_as_sf(coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))


# LOAD IN THE COUNTRY AND REGION MAP DATA ----

country <- st_read(here("data", "spatial", "ghana_boundaries_shp", "gha_admbnda_adm0_gss_20210308.shp"), quiet = TRUE)

regions <- st_read(here("data", "spatial", "ghana_boundaries_shp",  "gha_admbnda_adm1_gss_20210308.shp"), quiet = TRUE) %>%
  rename(region = ADM1_EN)

bono_east <- regions %>% filter(region == "Bono East")


## SPECIFIC TOWN MARKERS ----

# just the lat/lon for kintampo
kintampo_sf <- st_as_sf(data.frame(
  location = "Kintampo",
  geo_lon = -1.7296,
  geo_lat = 8.0593
), coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))

# Dwenewoho
dwenewoho_sf <- st_as_sf(data.frame(
  location = "dwenewoho",
  geo_lon = -1.8325,
  geo_lat = 7.7403
), coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))


# New Longoro
new_longoro_sf <- st_as_sf(data.frame(
  location = "New Longoro",
  geo_lon = -2.02953,
  geo_lat = 8.14399
), coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))

# Apesika
apesika_sf <- st_as_sf(data.frame(
  location = "Apesika",
  geo_lon = -1.5453,
  geo_lat = 7.99838
), coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))

# Kurawura Akura
kurawura_akura_sf <- st_as_sf(data.frame(
  location = "Kurawura Akura",
  geo_lon = -1.4684,
  geo_lat = 8.7380
), coords = c("geo_lon", "geo_lat"), crs = st_crs(4326))


# LOAD IN THE ROADS DATA ----

# Load roads data and set CRS (assuming it's EPSG:4326)
roads <- st_read(here("data", "spatial", "ghana_roads_shp", "hotosm_gha_roads_lines_shp.shp"), quiet = TRUE)
st_crs(roads) <- 4326

#Filter the roads to only bono east
roads_filtered <- st_intersection(roads, bono_east)
```


## Monitor Locations

```{r plot monitor locations, echo=FALSE, message=FALSE, warning=FALSE}
# Define the bounding box coordinates
bbox_coords <- matrix(c(-2.2, 7.6, -2.2, 8.8, -1.3, 8.8, -1.3, 7.6, -2.2, 7.6), ncol = 2, byrow = TRUE)
bbox_polygon <- st_polygon(list(bbox_coords))
bbox_sf <- st_sfc(bbox_polygon, crs = st_crs(4326))

# Plot the map with the bounding box
ggplot() +
  # Add Ghana regions
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5) +
  # Add monitor points with transparency
  geom_sf(data = monitor_points, aes(geometry = geometry), color = "red", alpha = 0.5, size = 2) +
  # Add bounding box
  geom_sf(data = bbox_sf, fill = NA, color = "blue", lwd = 0.5, linetype = "solid") +
  # Add Ghana country boundary
  geom_sf(data = country, fill = NA, color = "black", size = 1) +
  #labs(title = "Monitor Locations in Ghana") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 1), # Increase tick mark size
    plot.title = element_text(size = 16) 
  )


# Save as high resolution PNG
ggsave("monitor_map.png", monitor_map, width = 8, height = 6, dpi = 300)



monitor_map_zoomed <- ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#ffe493", "#dadbe0"), color = "black", alpha = 0.5) +
  # Add monitor points with transparency within the bounding box
  geom_sf(data = monitor_points %>%
          mutate(geometry = st_jitter(geometry, amount = 0.001)), 
        aes(color = monitor_type), alpha = 0.6, size = 5) +
  
  # Add Kintampo marker
  geom_sf(data = kintampo_sf, aes(geometry = geometry), shape = 20, size = 3, color = "black") +
  # Add Dwenewoho marker
  geom_sf(data = dwenewoho_sf, aes(geometry = geometry), shape = 20, size = 3, color = "black") +
  # Add new longoro marker
  geom_sf(data = new_longoro_sf, aes(geometry = geometry), shape = 20, size = 3, color = "black") +
  # Add apesika marker
  geom_sf(data = apesika_sf, aes(geometry = geometry), shape = 20, size = 3, color = "black") +
  # Add kurawura akura marker
  geom_sf(data = kurawura_akura_sf, aes(geometry = geometry), shape = 20, size = 3, color = "black") +
  
  # Add Kintamp annotation text
  annotate("text", x = -1.73, y = 8.10, label = "Kintampo", color = "black", size = 5) +
  # Add Dwenewoho annotation text
  annotate("text", x = -1.69, y = 7.74, label = "Dwenewoho", color = "black", size = 5) +
  # Add new longoro annotation text
  annotate("text", x = -2.047, y = 8.214, label = "New Longoro", color = "black", size = 5) +
  # Add apesika annotation text
  annotate("text", x = -1.459, y = 7.96, label = "Apesika", color = "black", size = 5) +
  # Add atta akura annotation text
  annotate("text", x = -1.65, y = 8.75, label = "Kurawura Akura", color = "black", size = 5) +
  
  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  labs(x = "", 
       y = "",
       color = "Monitor Type") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.5), # Increase tick mark size
    plot.title = element_text(size = 16), # Increase title size
    legend.text = element_text(size = 10), # Increase legend text size
    legend.title = element_text(size = 12) # Increase legend title size
  )

# Save as high resolution PNG
ggsave("monitor_map_zoomed.png", monitor_map_zoomed, width = 8, height = 6, dpi = 300)
```


## Load corrected pollution data and create summaries 

```{r}
# LOAD CORRECTED FULL POLLUTION DATA ----
pm1_corrected <- readRDS(here("data", "pm", "final", "pm1corrected_20231024-20240816.rds"))
pm25_corrected <- readRDS(here("data", "pm", "final", "pm25corrected_20231024-20240816.rds"))
pm10_corrected <- readRDS(here("data", "pm", "final", "pm10corrected_20231024-20240816.rds"))

## LOAD CORRECTED SUMMARIZED POLLUTION DATA ----
pm1_community_hourly <- readRDS(here("data", "pm", "summarized", "pm1_community_hourly_20231024-20240816.rds"))
pm1_community_daily <- readRDS(here("data", "pm", "summarized", "pm1_community_daily_20231024-20240816.rds"))

pm25_community_hourly <- readRDS(here("data", "pm", "summarized", "pm25_community_hourly_20231024-20240816.rds"))
pm25_community_daily <- readRDS(here("data", "pm", "summarized", "pm25_community_daily_20231024-20240816.rds"))

pm10_community_hourly <- readRDS(here("data", "pm", "summarized", "pm10_community_hourly_20231024-20240816.rds"))
pm10_community_daily <- readRDS(here("data", "pm", "summarized", "pm10_community_daily_20231024-20240816.rds"))
```


```{r}
#monitor missingness
monitor_missingness <- pm25_corrected %>% 
  group_by(monitor) %>% 
  summarise(missing_rate = mean(is.na(pm25)) * 100) %>%
  mutate(monitor = fct_reorder(monitor, -missing_rate))

# MAP OF MONITOR MISSINGNESS BY LOCATION ----
missing_monitor_location <- left_join(monitor_missingness, monitor_points)
```


## Data source (Cloud vs SD for each monitor)

```{r plot cloud vs sd data, echo=FALSE, message=FALSE, warning=FALSE}
monitor_obs_count_type <- pm25_corrected %>%
  group_by(monitor, source) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  group_by(monitor) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ungroup() %>%
  filter(!is.na(source)) %>%
  mutate(source = factor(source, levels = c("sd_card", "cloud")))

monitor_order <- monitor_obs_count_type %>%
  filter(source %in% c("cloud", "sd_card")) %>%
  group_by(monitor) %>%
  summarise(total = sum(percentage)) %>%
  arrange(desc(total)) %>%
  pull(monitor)

monitor_obs_count_type <- monitor_obs_count_type %>%
  mutate(monitor = factor(monitor, levels = monitor_order))

ggplot(monitor_obs_count_type, aes(x = monitor, y = percentage, fill = source)) +
  geom_bar(stat = "identity") +
  labs(title = "",
       x = "Monitor",
       y = "Percentage of Data Recorded",
       fill = "Data Source") +  # Custom legend title
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_text(size = 14),  # Increase x-axis title size
        axis.title.y = element_text(size = 14),  # Increase y-axis title size
        legend.text = element_text(size = 12),   # Increase legend text size
        legend.title = element_text(size = 14)) +  # Increase legend title size
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  scale_fill_manual(values = c("cloud" = "#2599db", "sd_card" = "#b54033"),
                    labels = c("cloud" = "Cloud", "sd_card" = "SD Card"))


```



# Pollutant Trends

## Grand Average plots 

```{r grand average facet, echo=FALSE, message=FALSE, warning=FALSE}
# join daily PM into one dataset
daily_pollutants <- left_join(pm1_community_daily, pm25_community_daily) %>% 
  left_join(pm10_community_daily)

daily_pollutants$date <- as.Date(daily_pollutants$date)


# pivot longer data so each measurement is its own row 
daily_pollutants_long <- daily_pollutants %>%
  group_by(date) %>%
  summarise('Mean PM 1' = mean(mean_pm1, na.rm = TRUE),
            'Mean PM 2.5' = mean(mean_pm25, na.rm = TRUE),
            'Mean PM 10' = mean(mean_pm10, na.rm = TRUE)) %>%
  ungroup() %>%
  pivot_longer(cols = c("Mean PM 1", "Mean PM 2.5", "Mean PM 10"), names_to = "pollutant", values_to = "measurement")

# Calculate the mean values for each pollutant
mean_values <- daily_pollutants_long %>%
  group_by(pollutant) %>%
  summarise(mean_measurement = signif(mean(measurement, na.rm = TRUE), 3))

# Merge the mean values back into the original data frame for plotting
daily_pollutants_long <- daily_pollutants_long %>%
  left_join(mean_values, by = "pollutant")


# Convert the 'pollutant' column to a factor with the desired order
daily_pollutants_long$pollutant <- factor(daily_pollutants_long$pollutant, levels = c("Mean PM 1", "Mean PM 2.5", "Mean PM 10"))


facet_pm_time_series <- daily_pollutants_long %>%
  ggplot(aes(x = date, y = measurement)) +
  geom_line() +
  geom_hline(aes(yintercept = mean_measurement), color = "#cc1414", linetype = "dashed") +
  geom_text(data = mean_values, aes(x = max(daily_pollutants_long$date), y = mean_measurement, label = paste0(pollutant, ": ", round(mean_measurement, 2), " µg/m³")),
            color = "#cc1414", hjust = 1, vjust = -0.9, size = 4.5) +
  theme_bw() +
  labs(x = "Date",
       y = "Pollutant Measurement (µg/m³)") +
  facet_grid(factor(pollutant, levels = c("Mean PM 1", "Mean PM 2.5", "Mean PM 10" )) ~ ., scales = "free_y") +
  theme(
    axis.title = element_text(size = 14),   # Increase axis titles size
    axis.text = element_text(size = 12),    # Increase axis labels size
    strip.text = element_text(size = 12)  # Increase facet group text size
  )


# Save as high resolution PNG
ggsave("facet_pm_time_series.png", facet_pm_time_series, width = 8, height = 6, dpi = 300)
```


```{r pm25 aqi plot, echo=FALSE, message=FALSE, warning=FALSE}

# Define AQI levels with ordered factors
aqi_levels <- data.frame(
  ymin = c(0, 9.1, 35.5, 55.5, 125.5, 225.5),
  ymax = c(9, 35.4, 55.4, 125.4, 225.4, 450),
  fill = factor(c("Good", "Moderate", "Unhealthy for Sensitive Groups", 
                  "Unhealthy", "Very Unhealthy", "Hazardous"), 
                levels = c("Good", "Moderate", "Unhealthy for Sensitive Groups", 
                           "Unhealthy", "Very Unhealthy", "Hazardous")),
  color = c("green", "yellow", "orange", "red", "purple", "maroon")
)

# Plot with legend inside the graph
pm25_colors <- daily_pollutants_long %>%
  filter(pollutant == "Mean PM 2.5") %>% 
  ggplot(aes(x = date, y = measurement)) +
  # Add background rectangles for AQI levels
  geom_rect(data = aqi_levels, 
            aes(xmin = min(daily_pollutants_long$date), 
                xmax = max(daily_pollutants_long$date), 
                ymin = ymin, 
                ymax = ymax, 
                fill = fill), 
            color = NA, alpha = 0.2, inherit.aes = FALSE) +
  # Add the PM 2.5 line
  geom_line(color = "black") +
  theme_bw() +
  labs(x = "Date",
       y = bquote("Fleet Mean " ~ PM[2.5] ~ (µg/m^3)),
       fill = "US-EPA AQI Risk Categories") +  # Add legend title for the AQI levels
  theme(
    axis.title = element_text(size = 14),   # Increase axis titles size
    axis.text = element_text(size = 12),    # Increase axis labels size
    strip.text = element_text(size = 12),    # Increase facet group text size
    legend.position = c(0.77, 0.73),  # Position legend inside the graph (top right)
    legend.background = element_rect(fill = rgb(1, 1, 1, alpha = 0.7), color = "black"),
    legend.title = element_text(size = 12),  # Adjust legend title size
    legend.text = element_text(size = 10)    # Adjust legend text size
  ) +
  scale_fill_manual(values = aqi_levels$color) +
  guides(fill = guide_legend(reverse = TRUE))  # Reverse the order of the legend


ggsave("pm25_colors_time_series.png", pm25_colors, width = 8, height = 4, dpi = 300)


```


```{r pm25 aqi vertical stripes plot, echo=FALSE, message=FALSE, warning=FALSE}

# First, categorize the measurement values into AQI levels
daily_pollutants_long <- daily_pollutants_long %>%
  mutate(AQI_Category = cut(measurement, 
                            breaks = c(-Inf, 9, 35.4, 55.4, 125.4, 225.4, Inf), 
                            labels = c("Good", "Moderate", "Unhealthy for Sensitive Groups", 
                                       "Unhealthy", "Very Unhealthy", "Hazardous"),
                            right = FALSE))

# Plot with vertical bands and legend on the side
pm25_vertical_stripes <- ggplot(daily_pollutants_long %>% filter(pollutant == "Mean PM 2.5"), 
       aes(x = date, y = measurement)) +
  # Add background tiles for AQI levels
  geom_tile(aes(fill = AQI_Category), height = Inf, alpha = 0.3) +
  # Add the PM 2.5 line
  geom_line(color = "black") +
  theme_bw() +
  labs(x = "Date",
       y = bquote("Fleet Mean " ~ PM[2.5] ~ (µg/m^3)),
       fill = "US-EPA AQI Risk Categories") +  # Add legend title for the AQI levels
  theme(
    axis.title = element_text(size = 14),   # Increase axis titles size
    axis.text = element_text(size = 12),    # Increase axis labels size
    strip.text = element_text(size = 12),    # Increase facet group text size
    legend.position = "right",  # Position legend on the right side
    legend.title = element_text(size = 12),  # Adjust legend title size
    legend.text = element_text(size = 10)    # Adjust legend text size
  ) +
  scale_fill_manual(values = aqi_levels$color) +
  guides(fill = guide_legend(reverse = TRUE))  # Reverse the order of the legend


ggsave("pm25_vertical_stripes.png", pm25_vertical_stripes, width = 8, height = 4, dpi = 300)


```

## PM 2.5 Community Deployment

### Checking Individual Monitor to Fleet Average

```{r monitor vs fleet avg 25, fig.height=10, fig.width=7, echo=FALSE, message=FALSE, warning=FALSE}
# hourly comparison
compare_fleet_avg_monitor(pm25_community_hourly, "pm25", "hourly")

# Daily comparison
compare_fleet_avg_monitor(pm25_community_daily, "pm25", "daily")
```

#### PM 2.5 Hourly: Individual Monitor Regressed on Fleet Average
**Note: fleet average required at minimum 5 readings to be included in regression)**
```{r monitor vs fleet avg 25 hourly table, echo=FALSE, message=FALSE, warning=FALSE}
pm25_hourly_regression_data <- pm25_community_hourly %>% 
  group_by(date, hour) %>% 
  summarize(offline_monitors = sum(is.na(mean_pm25))) %>%
  ungroup() %>%
  
  #add the PM data back
  left_join(pm25_community_hourly) %>%
  
  filter(offline_monitors < 35) %>%
  
  filter(!is.na(mean_pm25))


pm25_community_regression_results <- apply_regression(pm25_hourly_regression_data, "mean_pm25", "fleet_average_pm25")

datatable(pm25_community_regression_results)
```

#### PM 2.5 Daily: Individual Monitor Regressed on Fleet Average
**Note: fleet average required at minimum 5 readings to be included in regression)**
```{r monitor vs fleet avg 25 daily table, echo=FALSE, message=FALSE, warning=FALSE}
pm25_daily_regression_data <- pm25_community_daily %>% 
  group_by(date) %>% 
  summarize(offline_monitors = sum(is.na(mean_pm25))) %>%
  ungroup() %>%
  
  #add the PM data back
  left_join(pm25_community_daily) %>%
  
  filter(offline_monitors < 35) %>%
  
  filter(!is.na(mean_pm25))


pm25_community_regression_results <- apply_regression(pm25_daily_regression_data, "mean_pm25", "fleet_average_pm25")

datatable(pm25_community_regression_results)
```


### PM 2.5 Trends

```{r pm25 timelines, echo=FALSE, message=FALSE, warning=FALSE}
daily_pm25_all_devices <- pm25_corrected %>%
  group_by(date, monitor) %>%
  summarize(mean_pm25 = mean(pm25, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_pm25, group = monitor)) +
  geom_line(color = "black", alpha = 0.2) +
  theme_minimal() +
  labs(y = "Mean PM 2.5",
       title = "Average PM 2.5 Accross All Devices")

daily_cooktime_pm25_all_devices <- pm25_corrected %>%
  filter(hour >= 16 & hour < 20) %>%
  group_by(monitor, date) %>%
  summarise(mean_pm25 = mean(pm25, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_pm25, group = monitor)) +
  geom_line(color = "black", alpha = 0.4) +
  theme_minimal() +
  labs(y = "Mean PM 2.5",
       title = "Daily Average PM 2.5 for Each Monitor During Primary Cooking Hours (4-8)") 

grid.arrange(daily_pm25_all_devices, daily_cooktime_pm25_all_devices)

daily_pm25_timeline <- pm25_corrected %>%
  group_by(date) %>%
  summarize(mean_pm25 = mean(pm25, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_pm25)) +
  geom_line() +
  theme_minimal() +
  labs(y = "Mean PM 2.5",
       title = "Average PM 2.5 Accross All Devices")

daily_cooktime_pm25_timeline <- pm25_corrected %>%
  filter(hour >= 16 & hour < 20) %>%
  group_by(date) %>%
  summarize(mean_pm25 = mean(pm25, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_pm25)) +
  geom_line() +
  theme_minimal() +
  labs(y = "Mean PM 2.5",
       title = "Mean PM 2.5 Accross All Devices During Cooking Hours (4-8)")

grid.arrange(daily_pm25_timeline, daily_cooktime_pm25_timeline)
```

### PM 2.5 Heatmap

```{r pm25 trends, echo=FALSE, message=FALSE, warning=FALSE}
generate_heatmap(pm25_community_hourly, "pm25")
```

### HEATMAP CUSTOMIZATION

```{r heatmap customization, echo = FALSE, warning=FALSE, message=FALSE}
## PM 25
 
pm25_community_hourly <- pm25_community_hourly %>%
    mutate(date = ymd(date))  # Use lubridate to ensure date conversion
  
heat_map_data <- pm25_community_hourly %>%
    group_by(date, hour) %>%
    summarize(mean_pm25 = mean(mean_pm25, na.rm = TRUE), .groups = 'drop') %>%
    ungroup() %>%
    mutate(day = day(date),
           month = factor(format(date, "%b"), levels = c("Sep", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug")),
           year = year(date)) %>%
  mutate(log_mean_pm25 = log(mean_pm25))


pm_heatmap <- ggplot(heat_map_data, aes(day, hour, fill = mean_pm25)) +
    geom_tile(color = NA, size = 0) + 
    scale_fill_gradientn(
        colors = c("#221f26", "#3f2363", "#872CA2", "#D44292", "#F98477", "#EDD9A1", "#fffceb"), # Specify your custom colors
        values = scales::rescale(c(0, 5, 12, 20, 40, 60, 100)), # Set breaks; adjust these values based on your data range
        name = "PM 2.5 \n(µg/m³)"
    ) + 
    facet_grid(~month) +
    scale_y_continuous(trans = "reverse", breaks = unique(heat_map_data$hour)) +
    scale_x_continuous(breaks = c(1, 10, 20, 31)) +
    #theme_minimal(base_size = 8) +
    labs(x = "Day",
         y = "Hour") + 
    theme(legend.position = "right",
          plot.title = element_text(size = 14, hjust = 0),
          axis.text.y = element_text(size = 10),
          strip.background = element_rect(colour = "white"),
          strip.text = element_text(size = 13),   # Increase facet group text size
          axis.ticks = element_blank(),
          axis.text = element_text(size = 10),
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 11),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14),
          panel.background = element_blank())

ggsave("pm25_heatmap.png", pm_heatmap, width = 10, height = 4, dpi = 300)



## PM 10 ----

pm10_community_hourly <- pm10_community_hourly %>%
    mutate(date = ymd(date))  # Use lubridate to ensure date conversion
  
heat_map_data <- pm10_community_hourly %>%
    group_by(date, hour) %>%
    summarize(mean_pm10 = mean(mean_pm10, na.rm = TRUE), .groups = 'drop') %>%
    ungroup() %>%
    mutate(day = day(date),
           month = factor(format(date, "%b"), levels = c("Sep", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug")),
           year = year(date)) %>%
  mutate(log_mean_pm10 = log(mean_pm10))


pm10_heatmap <- ggplot(heat_map_data, aes(day, hour, fill = mean_pm10)) +
    geom_tile(color = NA, size = 0) + 
    scale_fill_gradientn(
        colors = c("#221f26", "#3f2363", "#872CA2", "#D44292", "#F98477", "#EDD9A1", "#fffceb"), 
        values = scales::rescale(c(0, 5, 10, 20, 40, 60, 100)),
        breaks = c(5, 250, 500, 750, 950),  # Add more breaks
        labels = c("5", "250", "500", "750", "950"),  # Corresponding labels
        name = "PM 10 \n(µg/m³)"
    ) + 
    facet_grid(~month) +
    scale_y_continuous(trans = "reverse", breaks = unique(heat_map_data$hour)) +
    scale_x_continuous(breaks = c(1, 10, 20, 31)) +
    #theme_minimal(base_size = 8) +
    labs(x = "Day",
         y = "Hour") + 
    theme(legend.position = "right",
          plot.title = element_text(size = 14, hjust = 0),
          axis.text.y = element_text(size = 10),
          strip.background = element_rect(colour = "white"),
          strip.text = element_text(size = 13),  
          axis.ticks = element_blank(),
          axis.text = element_text(size = 10),
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 11),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14),
          panel.background = element_blank())

ggsave("pm10_heatmap.png", pm10_heatmap, width = 10, height = 4, dpi = 300)


```





### Harmattan vs non-Harmattan + Cooktime vs non-Cooktime

```{r pm25 trends harmattan and cooktime calc, echo=FALSE, message=FALSE, warning=FALSE}
harmattan_start <- as.Date("2023-12-01")
harmattan_end <- as.Date("2024-03-01")

pm25_community_hourly_seasonality <- pm25_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 15 & hour <= 18 ~ "Active Cooking Hours",
      (hour < 15 | hour > 18) & hour != 3 ~ "Non-Active Cooking Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Active Cooking Hours", "Non-Active Cooking Hours", "Control Hour"))
  )

# hourly pm 2.5 data from just harmattan only
harmattan_pm25_hourly <- pm25_community_hourly_seasonality %>% filter(date >= harmattan_start & date <= harmattan_end) %>% 
  filter(!is.na(hour)) 

# hourly pm 2.5 data from not harmattan only
not_harmattan_pm25_hourly <- pm25_community_hourly_seasonality %>% filter(date < harmattan_start | date > harmattan_end) %>%
  filter(!is.na(hour)) 

# mean and sd pm25 during harmattan
mean(harmattan_pm25_hourly$mean_pm25, na.rm = TRUE)
sd(harmattan_pm25_hourly$mean_pm25, na.rm = TRUE)

# mean and sd pm25 NOT during harmattan
mean(not_harmattan_pm25_hourly$mean_pm25, na.rm = TRUE)
sd(not_harmattan_pm25_hourly$mean_pm25, na.rm = TRUE)


# COMPARING COOKTIME PM 2.5 ACROSS ALL SEASONS ----
anova_result <- aov(mean_pm25 ~ cooking_period, data = pm25_community_hourly_seasonality)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)

# COMPARING COOKTIME PM 2.5 DURING HARMATTAN ----

anova_result <- aov(mean_pm25 ~ cooking_period, data = harmattan_pm25_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)

# COMPARING COOKTIME PM 2.5 NOT HARMATTAN ----

anova_result <- aov(mean_pm25 ~ cooking_period, data = not_harmattan_pm25_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)

# COOKTIME VS NOT COOKTIME COMP (MEANS AND SDs) -----

cooktime_pm25_hourly <- pm25_community_hourly_seasonality %>% filter(cooking_period == "Active Cooking Hours") %>% 
  filter(!is.na(hour)) 

not_cooktime_pm25_hourly <- pm25_community_hourly_seasonality %>% filter(cooking_period == "Non-Active Cooking Hours") %>% 
  filter(!is.na(hour)) 

control_pm25_hourly <- pm25_community_hourly_seasonality %>% filter(cooking_period == "Control Hour") %>% 
  filter(!is.na(hour)) 

mean(cooktime_pm25_hourly$mean_pm25, na.rm = TRUE) 
sd(cooktime_pm25_hourly$mean_pm25, na.rm = TRUE) 

mean(not_cooktime_pm25_hourly$mean_pm25, na.rm = TRUE)  
sd(not_cooktime_pm25_hourly$mean_pm25, na.rm = TRUE) 

mean(control_pm25_hourly$mean_pm25, na.rm = TRUE)  
sd(control_pm25_hourly$mean_pm25, na.rm = TRUE)

# COOKTIME VS NOT COOKTIME VS CONTROL DURING HARMATTAN (MEANS AND SDs) ----

mean(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Active Cooking Hours"], na.rm = TRUE)
sd(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Active Cooking Hours"], na.rm = TRUE)

mean(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Non-Active Cooking Hours"], na.rm = TRUE)
sd(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Non-Active Cooking Hours"], na.rm = TRUE)

mean(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Control Hour"], na.rm = TRUE)

# COOKTIME VS NOT COOKTIME NOT DURING HARMATTAN (MEANS AND SDs) ----

mean(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Active Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Active Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Non-Active Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Non-Active Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
```

```{r pm25 trends harmattan and cooktime plot, echo=FALSE, message=FALSE, warning=FALSE}
## PLOTTING THE COMPARISONS 

base_plot <- pm25_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 15 & hour <= 18 ~ "Active Cooking Hours",
      (hour < 15 | hour > 18) & hour != 3 ~ "Non-Active Cooking Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Active Cooking Hours", "Non-Active Cooking Hours", "Control Hour"))
  ) %>%
  filter(!is.na(hour)) %>%
  ggplot(aes(x = harmattan, y = mean_pm25, fill = cooking_period)) +
  geom_boxplot() +
  theme_classic() +
  labs(y = "PM 2.5 Concentration", 
       x = "Season",
       fill = "Cooking\nPeriod") +  # Change legend title and add line breaks
  theme(legend.position = "right") +
  scale_fill_manual(values = c(
    "Active Cooking Hours" = "goldenrod", 
    "Non-Active Cooking Hours" = "#52b3eb", 
    "Control Hour" = "#b672e0"
  ),
  labels = c("Evening Cooking Hours (4pm-8pm)", "Other Hours (Excluding Evening Cooking \nand Control Hours)", "Control Hour (3am)"))  # Add line breaks in legend labels


## NOTE: USED RESULTS FROM ANOVA AND TUKEY TESTS TO DETERMINE SIGNFICANCE STARTS. WILL NEED TO BE UPDATED DEPENEDING ON NEW DATA. 
base_plot +
  geom_segment(aes(x = 0.72, xend = 0.98, y = 545, yend = 545), color = "black") +  # Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 0.85, y = 550, label = "***") +
  
  geom_segment(aes(x = 1.02, xend = 1.28, y = 545, yend = 545), color = "black") +  # Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 1.15, y = 550, label = "***") +
  
  geom_segment(aes(x = 0.72, xend = 1.27, y = 565, yend = 565), color = "black") +  # Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 1, y = 570, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 1.98, y = 360, yend = 360), color = "black") +  # Not Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 1.85, y = 365, label = "***") +
  
  geom_segment(aes(x = 2.02, xend = 2.28, y = 360, yend = 360), color = "black") +  # Not Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 2.15, y = 365, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 2.27, y = 380, yend = 380), color = "black") +  # Not Harmattan Cooking Hour vs Not Cooking Hour
  annotate("text", x = 2, y = 385, label = "***") +
  
  geom_segment(aes(x = 1, xend = 2, y = 590, yend = 590), color = "black") +  # Harmattan vs Not Harmattan
  annotate("text", x = 1.5, y = 600, label = "***") +
  
  labs(caption ="Comparisons and their significance are denoted by the horizontal lines, where the three stars indicate p < 0.001.") +
  
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    legend.text = element_text(size = 10), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    plot.caption = element_text(size = 11),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
  )
```

### Harmattan vs non-Harmattan + Cooktime vs non-Cooktime (evening vs morning cooking hours)


```{r pm25 trends harmattan and cooktime calc, echo=FALSE, message=FALSE, warning=FALSE}
harmattan_start <- as.Date("2023-12-01")
harmattan_end <- as.Date("2024-03-01")

pm25_community_hourly_seasonality <- pm25_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 16 & hour <= 19 ~ "Evening Cooking Hours", #4-9pm
      hour >= 5 & hour <= 8 ~ "Morning Cooking Hours", #5-9am
      (hour < 16 | hour > 19) &  (hour < 5 | hour > 8) & hour != 3 ~ "Other Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Evening Cooking Hours", "Morning Cooking Hours", "Control Hour", "Other Hours"))
  )

# hourly pm 2.5 data from just harmattan only
harmattan_pm25_hourly <- pm25_community_hourly_seasonality %>% filter(date >= harmattan_start & date <= harmattan_end) %>% 
  filter(!is.na(hour)) 

# hourly pm 2.5 data from not harmattan only
not_harmattan_pm25_hourly <- pm25_community_hourly_seasonality %>% filter(date < harmattan_start | date > harmattan_end) %>%
  filter(!is.na(hour)) 

# mean and sd pm25 during harmattan
mean(harmattan_pm25_hourly$mean_pm25, na.rm = TRUE)
sd(harmattan_pm25_hourly$mean_pm25, na.rm = TRUE)

# mean and sd pm25 NOT during harmattan
mean(not_harmattan_pm25_hourly$mean_pm25, na.rm = TRUE)
sd(not_harmattan_pm25_hourly$mean_pm25, na.rm = TRUE)

# COMPARING COOKTIME PM 2.5 ACROSS ALL SEASONS ---- 
anova_result <- aov(mean_pm25 ~ cooking_period, data = pm25_community_hourly_seasonality)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)

# COMPARING COOKTIME PM 2.5 DURING HARMATTAN ---- 

anova_result <- aov(mean_pm25 ~ cooking_period, data = harmattan_pm25_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)

# COMPARING COOKTIME PM 2.5 NOT HARMATTAN ----

anova_result <- aov(mean_pm25 ~ cooking_period, data = not_harmattan_pm25_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)

# COOKTIME VS NOT COOKTIME COMP (MEANS AND SDs) ----

evening_cooktime_pm25_hourly <- pm25_community_hourly_seasonality %>% filter(cooking_period == "Evening Cooking Hours") %>% 
  filter(!is.na(hour)) 

morning_cooktime_pm25_hourly <- pm25_community_hourly_seasonality %>% filter(cooking_period == "Morning Cooking Hours") %>% 
  filter(!is.na(hour)) 

other_pm25_hourly <- pm25_community_hourly_seasonality %>% filter(cooking_period == "Other Hours") %>% 
  filter(!is.na(hour)) 

control_pm25_hourly <- pm25_community_hourly_seasonality %>% filter(cooking_period == "Control Hour") %>% 
  filter(!is.na(hour)) 

mean(evening_cooktime_pm25_hourly$mean_pm25, na.rm = TRUE) 
sd(evening_cooktime_pm25_hourly$mean_pm25, na.rm = TRUE) 

mean(morning_cooktime_pm25_hourly$mean_pm25, na.rm = TRUE)  
sd(morning_cooktime_pm25_hourly$mean_pm25, na.rm = TRUE) 

mean(other_pm25_hourly$mean_pm25, na.rm = TRUE)  
sd(other_pm25_hourly$mean_pm25, na.rm = TRUE)

mean(control_pm25_hourly$mean_pm25, na.rm = TRUE)  
sd(control_pm25_hourly$mean_pm25, na.rm = TRUE)

# COOKTIME VS NOT COOKTIME VS CONTROL DURING HARMATTAN (MEANS AND SDs) ----

mean(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)
sd(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)

mean(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)
sd(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)

mean(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Other Hours"], na.rm = TRUE)
sd(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Other Hours"], na.rm = TRUE)

mean(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(harmattan_pm25_hourly$mean_pm25[harmattan_pm25_hourly$cooking_period == "Control Hour"], na.rm = TRUE)

# COOKTIME VS NOT COOKTIME NOT DURING HARMATTAN (MEANS AND SDs)

mean(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Other Hours"], na.rm = TRUE)
sd(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Other Hours"], na.rm = TRUE)

mean(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(not_harmattan_pm25_hourly$mean_pm25[not_harmattan_pm25_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
```

```{r pm25 trends harmattan and cooktime plot, echo=FALSE, message=FALSE, warning=FALSE}
## PLOTTING THE COMPARISONS 

base_plot <- pm25_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 16 & hour <= 19 ~ "Evening Cooking Hours", #4-9pm
      hour >= 5 & hour <= 8 ~ "Morning Cooking Hours", #5-9am
      (hour < 16 | hour > 19) &  (hour < 5 | hour > 8) & hour != 3 ~ "Other Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Evening Cooking Hours", "Morning Cooking Hours", "Control Hour", "Other Hours"))
  ) %>%
  filter(!is.na(hour)) %>%
  ggplot(aes(x = harmattan, y = mean_pm25, fill = cooking_period)) +
  geom_boxplot() +
  theme_classic() +
  labs(y = "PM 2.5 Concentration", 
       x = "Season",
       fill = "Cooking\nPeriod") +  # Change legend title and add line breaks
  theme(legend.position = "right") +
  scale_fill_manual(values = c(
    "Evening Cooking Hours" = "goldenrod", 
    "Morning Cooking Hours" = "#7db569", 
    "Control Hour" = "#b672e0",
    "Other Hours" = "#52b3eb"
  ),
  labels = c("Evening Cooking Hours (4pm-8pm)", "Morning Cooking Hours (5am-9am)", "Control Hour (3am)", "Other Hours (Excluding Cooking \nand Control Hours)"))  # Add line breaks in legend labels

## NOTE: USED RESULTS FROM ANOVA AND TUKEY TESTS TO DETERMINE SIGNFICANCE STARTS. WILL NEED TO BE UPDATED DEPENEDING ON NEW DATA. 
full_boxplot_pm25 <- base_plot +
  geom_segment(aes(x = 0.72, xend = 0.90, y = 515, yend = 515), color = "black") +  # Harmattan 
  annotate("text", x = 0.81, y = 518, label = "***") +
  
  geom_segment(aes(x = 0.91, xend = 1.09, y = 515, yend = 515), color = "black") +  # Harmattan 
  annotate("text", x = 1, y = 518, label = "***") +
  
  geom_segment(aes(x = 1.10, xend = 1.28, y = 515, yend = 515), color = "black") +  # Harmattan 
  annotate("text", x = 1.19, y = 518, label = "**") +
  
  geom_segment(aes(x = 0.72, xend = 1.09, y = 525, yend = 525), color = "black") +  # Harmattan 
  annotate("text", x = 0.9, y = 529, label = "***") +
  
  geom_segment(aes(x = 0.91, xend = 1.27, y = 536, yend = 536), color = "black") +  # Harmattan
  annotate("text", x = 1.09, y = 540, label = "***") +
  
  geom_segment(aes(x = 0.72, xend = 1.27, y = 548, yend = 548), color = "black") +  # Harmattan
  annotate("text", x = 1, y = 552, label = "***") +
  
  
  
  geom_segment(aes(x = 1.72, xend = 1.90, y = 360, yend = 360), color = "black") +  # Not Harmattan 
  annotate("text", x = 1.81, y = 364, label = "***") +
  
  geom_segment(aes(x = 1.91, xend = 2.09, y = 360, yend = 360), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.0, y = 364, label = "***") +
  
  geom_segment(aes(x = 2.10, xend = 2.28, y = 360, yend = 360), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.19, y = 364, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 2.09, y = 372, yend = 372), color = "black") +  # Not Harmattan 
  annotate("text", x = 1.9, y = 376, label = "***") +
  
  geom_segment(aes(x = 1.91, xend = 2.28, y = 384, yend = 384), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.09, y = 388, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 2.27, y = 396, yend = 396), color = "black") +  # Not Harmattan 
  annotate("text", x = 2, y = 400, label = "***") +
  
  
  geom_segment(aes(x = 1, xend = 2, y = 590, yend = 590), color = "black") +  # Harmattan vs Not Harmattan
  annotate("text", x = 1.5, y = 600, label = "***") +
  
  labs(caption ="Comparisons and their significance are denoted by the horizontal lines, where \n* indicates p <0.05, ** indicates p < 0.01, and *** indicates p < 0.001") +
  
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    legend.text = element_text(size = 10), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    plot.caption = element_text(size = 11),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
  )


ggsave("full_boxplot_pm25.png", full_boxplot_pm25, width = 10, height = 8, dpi = 300)



```


### PM10 Harmattan vs non-Harmattan + Cooktime vs non-Cooktime (evening vs morning cooking hours)


```{r pm10 trends harmattan and cooktime calc, echo=FALSE, message=FALSE, warning=FALSE}
harmattan_start <- as.Date("2023-12-01")
harmattan_end <- as.Date("2024-03-01")

pm10_community_hourly_seasonality <- pm10_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 16 & hour <= 19 ~ "Evening Cooking Hours", #4-9pm
      hour >= 5 & hour <= 8 ~ "Morning Cooking Hours", #5-9am
      (hour < 16 | hour > 19) &  (hour < 5 | hour > 8) & hour != 3 ~ "Other Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Evening Cooking Hours", "Morning Cooking Hours", "Control Hour", "Other Hours"))
  )

# hourly pm 2.5 data from just harmattan only
harmattan_pm10_hourly <- pm10_community_hourly_seasonality %>% filter(date >= harmattan_start & date <= harmattan_end) %>% 
  filter(!is.na(hour)) 

# hourly pm 2.5 data from not harmattan only
not_harmattan_pm10_hourly <- pm10_community_hourly_seasonality %>% filter(date < harmattan_start | date > harmattan_end) %>%
  filter(!is.na(hour)) 

# mean and sd pm10 during harmattan
mean(harmattan_pm10_hourly$mean_pm10, na.rm = TRUE)
sd(harmattan_pm10_hourly$mean_pm10, na.rm = TRUE)

# mean and sd pm10 NOT during harmattan
mean(not_harmattan_pm10_hourly$mean_pm10, na.rm = TRUE)
sd(not_harmattan_pm10_hourly$mean_pm10, na.rm = TRUE)

# COMPARING COOKTIME PM 2.5 ACROSS ALL SEASONS ----
anova_result <- aov(mean_pm10 ~ cooking_period, data = pm10_community_hourly_seasonality)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)

# COMPARING COOKTIME PM 2.5 DURING HARMATTAN ----

anova_result <- aov(mean_pm10 ~ cooking_period, data = harmattan_pm10_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)

# COMPARING COOKTIME PM 2.5 NOT HARMATTAN ----

anova_result <- aov(mean_pm10 ~ cooking_period, data = not_harmattan_pm10_hourly)
summary(anova_result)

# Tukey's Honest Significant Difference test
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)

# COOKTIME VS NOT COOKTIME COMP (MEANS AND SDs) ----

evening_cooktime_pm10_hourly <- pm10_community_hourly_seasonality %>% filter(cooking_period == "Evening Cooking Hours") %>% 
  filter(!is.na(hour)) 

morning_cooktime_pm10_hourly <- pm10_community_hourly_seasonality %>% filter(cooking_period == "Morning Cooking Hours") %>% 
  filter(!is.na(hour)) 

other_pm10_hourly <- pm10_community_hourly_seasonality %>% filter(cooking_period == "Other Hours") %>% 
  filter(!is.na(hour)) 

control_pm10_hourly <- pm10_community_hourly_seasonality %>% filter(cooking_period == "Control Hour") %>% 
  filter(!is.na(hour)) 

mean(evening_cooktime_pm10_hourly$mean_pm10, na.rm = TRUE) 
sd(evening_cooktime_pm10_hourly$mean_pm10, na.rm = TRUE) 

mean(morning_cooktime_pm10_hourly$mean_pm10, na.rm = TRUE)  
sd(morning_cooktime_pm10_hourly$mean_pm10, na.rm = TRUE) 

mean(other_pm10_hourly$mean_pm10, na.rm = TRUE)  
sd(other_pm10_hourly$mean_pm10, na.rm = TRUE)

mean(control_pm10_hourly$mean_pm10, na.rm = TRUE)  
sd(control_pm10_hourly$mean_pm10, na.rm = TRUE)

# COOKTIME VS NOT COOKTIME VS CONTROL DURING HARMATTAN (MEANS AND SDs) ----

mean(harmattan_pm10_hourly$mean_pm10[harmattan_pm10_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)
sd(harmattan_pm10_hourly$mean_pm10[harmattan_pm10_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)

mean(harmattan_pm10_hourly$mean_pm10[harmattan_pm10_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)
sd(harmattan_pm10_hourly$mean_pm10[harmattan_pm10_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)

mean(harmattan_pm10_hourly$mean_pm10[harmattan_pm10_hourly$cooking_period == "Other Hours"], na.rm = TRUE)
sd(harmattan_pm10_hourly$mean_pm10[harmattan_pm10_hourly$cooking_period == "Other Hours"], na.rm = TRUE)

mean(harmattan_pm10_hourly$mean_pm10[harmattan_pm10_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(harmattan_pm10_hourly$mean_pm10[harmattan_pm10_hourly$cooking_period == "Control Hour"], na.rm = TRUE)

# COOKTIME VS NOT COOKTIME NOT DURING HARMATTAN (MEANS AND SDs)

mean(not_harmattan_pm10_hourly$mean_pm10[not_harmattan_pm10_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_pm10_hourly$mean_pm10[not_harmattan_pm10_hourly$cooking_period == "Evening Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_pm10_hourly$mean_pm10[not_harmattan_pm10_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)
sd(not_harmattan_pm10_hourly$mean_pm10[not_harmattan_pm10_hourly$cooking_period == "Morning Cooking Hours"], na.rm = TRUE)

mean(not_harmattan_pm10_hourly$mean_pm10[not_harmattan_pm10_hourly$cooking_period == "Other Hours"], na.rm = TRUE)
sd(not_harmattan_pm10_hourly$mean_pm10[not_harmattan_pm10_hourly$cooking_period == "Other Hours"], na.rm = TRUE)

mean(not_harmattan_pm10_hourly$mean_pm10[not_harmattan_pm10_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
sd(not_harmattan_pm10_hourly$mean_pm10[not_harmattan_pm10_hourly$cooking_period == "Control Hour"], na.rm = TRUE)
```

```{r pm10 trends harmattan and cooktime plot, echo=FALSE, message=FALSE, warning=FALSE}
## PLOTTING THE COMPARISONS ---- 
base_plot <- pm10_community_hourly %>% 
  mutate(harmattan = ifelse(date >= harmattan_start & date <= harmattan_end, "Harmattan", "Not Harmattan")) %>%
  mutate(
    cooking_period = factor(case_when(
      hour >= 16 & hour <= 19 ~ "Evening Cooking Hours", #4-9pm
      hour >= 5 & hour <= 8 ~ "Morning Cooking Hours", #5-9am
      (hour < 16 | hour > 19) &  (hour < 5 | hour > 8) & hour != 3 ~ "Other Hours",
      hour == 3 ~ "Control Hour"
    ), levels = c("Evening Cooking Hours", "Morning Cooking Hours", "Control Hour", "Other Hours"))
  ) %>%
  filter(!is.na(hour)) %>%
  ggplot(aes(x = harmattan, y = mean_pm10, fill = cooking_period)) +
  geom_boxplot() +
  theme_classic() +
  labs(y = "PM 10 Concentration", 
       x = "Season",
       fill = "Cooking\nPeriod") +  # Change legend title and add line breaks
  theme(legend.position = "right") +
  scale_fill_manual(values = c(
    "Evening Cooking Hours" = "goldenrod", 
    "Morning Cooking Hours" = "#7db569", 
    "Control Hour" = "#b672e0",
    "Other Hours" = "#52b3eb"
  ),
  labels = c("Evening Cooking Hours (4pm-8pm)", "Morning Cooking Hours (5am-9am)", "Control Hour (3am)", "Other Hours (Excluding Cooking \nand Control Hours)"))  # Add line breaks in legend labels

## NOTE: USED RESULTS FROM ANOVA AND TUKEY TESTS TO DETERMINE SIGNFICANCE STARTS. WILL NEED TO BE UPDATED DEPENEDING ON NEW DATA. 
full_boxplot_pm10 <- base_plot +
  geom_segment(aes(x = 0.72, xend = 0.90, y = 1400, yend = 1400), color = "black") +  # Harmattan 
  annotate("text", x = 0.81, y = 1410, label = "***") +
  
  geom_segment(aes(x = 0.91, xend = 1.09, y = 1400, yend = 1400), color = "black") +  # Harmattan 
  annotate("text", x = 1, y = 1410, label = "***") +
  
  geom_segment(aes(x = 1.10, xend = 1.28, y = 1400, yend = 1400), color = "black") +  # Harmattan 
  annotate("text", x = 1.19, y = 1410, label = "***") +
  
  geom_segment(aes(x = 0.72, xend = 1.09, y = 1440, yend = 1440), color = "black") +  # Harmattan 
  annotate("text", x = 0.9, y = 1450, label = "***") +
  
  geom_segment(aes(x = 0.91, xend = 1.27, y = 1480, yend = 1480), color = "black") +  # Harmattan
  annotate("text", x = 1.09, y = 1490, label = "***") +
  
  geom_segment(aes(x = 0.72, xend = 1.27, y = 1520, yend = 1520), color = "black") +  # Harmattan
  annotate("text", x = 1, y = 1530, label = "***") +
  
  
  
  geom_segment(aes(x = 1.72, xend = 1.90, y = 1150, yend = 1150), color = "black") +  # Not Harmattan 
  annotate("text", x = 1.81, y = 1160, label = "***") +
  
  geom_segment(aes(x = 1.91, xend = 2.09, y = 1150, yend = 1150), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.0, y = 1160, label = "***") +
  
  geom_segment(aes(x = 2.10, xend = 2.28, y = 1150, yend = 1150), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.19, y = 1160, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 2.09, y = 1190, yend = 1190), color = "black") +  # Not Harmattan 
  annotate("text", x = 1.9, y = 1200, label = "***") +
  
  geom_segment(aes(x = 1.91, xend = 2.28, y = 1230, yend = 1230), color = "black") +  # Not Harmattan 
  annotate("text", x = 2.09, y = 1240, label = "***") +
  
  geom_segment(aes(x = 1.72, xend = 2.27, y = 1270, yend = 1270), color = "black") +  # Not Harmattan 
  annotate("text", x = 2, y = 1280, label = "***") +
  
  
  geom_segment(aes(x = 1, xend = 2, y = 1560, yend = 1560), color = "black") +  # Harmattan vs Not Harmattan
  annotate("text", x = 1.5, y = 1575, label = "***") +
  
  labs(caption ="Comparisons and their significance are denoted by the horizontal lines, where \n* indicates p <0.05, ** indicates p < 0.01, and *** indicates p < 0.001") +
  
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    legend.text = element_text(size = 10), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    plot.caption = element_text(size = 11),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
  )



ggsave("full_boxplot_pm10.png", full_boxplot_pm10, width = 10, height = 8, dpi = 300)

```




### Spatial Correlation

```{r spatial correlation 25, echo=FALSE, message=FALSE, warning=FALSE}
generate_spatial_pollution_map(pm25_corrected, "pm25", missing_monitor_location)
```

### Calculate Moran I for Spatial Correlation

```{r moran i 25, echo=FALSE, message=FALSE, warning=FALSE}
calculate_moran_i(pm25_corrected, "pm25", missing_monitor_location)
```


## Moran I look nicer

```{r formatting moran i, echo=FALSE, warning=FALSE, message=FALSE}

# Define parameters
pollutant <- "pm25"
k <- 5

#file_name <- "pm25_moran_table.png"

## DURING HARMATTAN ----

# Calculate summary statistics
summary_stats <- pm25_corrected %>%
  filter(date >= harmattan_start & date < harmattan_end) %>%
  group_by(monitor) %>%
  summarize(
    mean_value = mean(!!sym(pollutant), na.rm = TRUE),
    median_value = median(!!sym(pollutant), na.rm = TRUE),
    sd_value = sd(!!sym(pollutant), na.rm = TRUE)
  )

monitor_data <- monitor_points %>%
  left_join(summary_stats, by = "monitor")

# Convert monitor data to spatial object using sf
monitor_sf <- st_as_sf(monitor_data)

# Calculate distance-based spatial weights
coords <- st_coordinates(monitor_sf)
nb <- knn2nb(knearneigh(coords, k = k)) # k-nearest neighbors
listw <- nb2listw(nb, style = "W")

# Calculate Moran's I
moran_result <- moran.test(monitor_sf$mean_value, listw)

# Extract relevant values directly
observed_value <- round(as.numeric(moran_result$estimate["Moran I statistic"]), 4)
expected_value <- round(as.numeric(moran_result$estimate["Expectation"]), 4)
variance <- round(as.numeric(moran_result$estimate["Variance"]), 4)
z_score <- round(as.numeric(moran_result$statistic), 4)
p_value <- round(as.numeric(moran_result$p.value), 4)

# Create the Moran's I results table
moran_table <- data.frame(
  Test = "Moran's I Statistic",
  `Observed Value` = observed_value,
  `Expected Value` = expected_value,
  `Variance` = variance,
  `Z-score` = z_score,
  `p-value` = p_value
)

# Format the table using kableExtra
formatted_table <- moran_table %>%
  kable("html", booktabs = TRUE, caption = "Moran's I Test Results: PM 2.5 Shows Spatial Correlation During Harmattan") %>%
  kable_styling(full_width = FALSE, position = "center", latex_options = c("hold_position", "striped"))



## OUTSIDE OF HARMATTAN ----

# Calculate summary statistics
summary_stats <- pm25_corrected %>%
  filter(date < harmattan_start | date <= harmattan_end) %>%
  group_by(monitor) %>%
  summarize(
    mean_value = mean(!!sym(pollutant), na.rm = TRUE),
    median_value = median(!!sym(pollutant), na.rm = TRUE),
    sd_value = sd(!!sym(pollutant), na.rm = TRUE)
  )

monitor_data <- monitor_points %>%
  left_join(summary_stats, by = "monitor")

# Convert monitor data to spatial object using sf
monitor_sf <- st_as_sf(monitor_data)

# Calculate distance-based spatial weights
coords <- st_coordinates(monitor_sf)
nb <- knn2nb(knearneigh(coords, k = k)) # k-nearest neighbors
listw <- nb2listw(nb, style = "W")

# Calculate Moran's I
moran_result <- moran.test(monitor_sf$mean_value, listw)

# Extract relevant values directly
observed_value <- round(as.numeric(moran_result$estimate["Moran I statistic"]), 4)
expected_value <- round(as.numeric(moran_result$estimate["Expectation"]), 4)
variance <- round(as.numeric(moran_result$estimate["Variance"]), 4)
z_score <- round(as.numeric(moran_result$statistic), 4)
p_value <- round(as.numeric(moran_result$p.value), 4)

# Create the Moran's I results table
moran_table <- data.frame(
  Test = "Moran's I Statistic",
  `Observed Value` = observed_value,
  `Expected Value` = expected_value,
  `Variance` = variance,
  `Z-score` = z_score,
  `p-value` = p_value
)

# Format the table using kableExtra
formatted_table <- moran_table %>%
  kable("html", booktabs = TRUE, caption = "Moran's I Test Results: PM 2.5 Shows Spatial Correlation Outside of Harmattan") %>%
  kable_styling(full_width = FALSE, position = "center", latex_options = c("hold_position", "striped"))

```


### Making maps for each size class

```{r pm maps during harmattan, echo=FALSE, message=FALSE, warning=FALSE}
# DURING HARMATTAN ----

#PM 2.5 ----
pm25_summary <- pm25_corrected %>%
  filter(date >= harmattan_start & date < harmattan_end) %>%
  group_by(monitor) %>%
  summarise(mean_pm25 = mean(pm25, na.rm = TRUE),
            median_pm25 = median(pm25, na.rm = TRUE)) %>%
  right_join(monitor_points)


ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5)  +
  
  # Add road data
  geom_sf(data = roads_filtered, color = "gray", size = 0.3) +  # Add roads data
  
  # Add monitor points with transparency within the bounding box
  geom_sf(data = pm25_summary, aes(geometry = geometry, color = mean_pm25), alpha = 0.7, size = 6) +

  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  
  # Customize
  labs(title = "Mean PM 2.5 for Each Monitor During Harmattan",
       x = "", 
       y = "") +
  theme_bw() +
  scale_color_gradient(low = "#0540a1", high = "#eb4c2d",
                       breaks = c(min(pm25_summary$mean_pm25), 75, 100, max(pm25_summary$mean_pm25)),
                       labels = c(paste(round(min(pm25_summary$mean_pm25),0)), "75", "100", paste(round(max(pm25_summary$mean_pm25),0))),
                       name = "PM 2.5 \n(µg/m³)") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    plot.title = element_text(size = 14), # Increase plot title size
    legend.text = element_text(size = 12), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    legend.key.size = unit(1.3, "lines") # Increase legend tick mark size
  )


#PM 1 ----
pm1_summary <- pm1_corrected %>%
  filter(date >= harmattan_start & date < harmattan_end) %>%
  group_by(monitor) %>%
  summarise(mean_pm1 = mean(pm1, na.rm = TRUE),
            median_pm1 = median(pm1, na.rm = TRUE)) %>%
  right_join(monitor_points)


ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5)  +
  
  # Add road data
  geom_sf(data = roads_filtered, color = "gray", size = 0.3) +  # Add roads data
  
  # Add monitor points with transparency within the bounding box
  geom_sf(data = pm1_summary, aes(geometry = geometry, color = mean_pm1), alpha = 0.7, size = 6) +

  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  
  # Customize
  labs(title = "Mean PM 1 for Each Monitor During Harmattan",
       x = "", 
       y = "") +
  theme_bw() +
  scale_color_gradient(low = "#0540a1", high = "#eb4c2d", 
                       breaks = c(min(pm1_summary$mean_pm1), 60, 85, max(pm1_summary$mean_pm1)),
                       labels = c(paste(round(min(pm1_summary$mean_pm1),0)), "60", "85", paste(round(max(pm1_summary$mean_pm1),0))),
                       name = "PM 1 \n(µg/m³)") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    plot.title = element_text(size = 14), # Incr ease plot title size
    legend.text = element_text(size = 12), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    legend.key.size = unit(1.3, "lines") # Increase legend tick mark size
  )


#PM 10 ----
pm10_summary <- pm10_corrected %>%
  filter(date >= harmattan_start & date < harmattan_end) %>%
  group_by(monitor) %>%
  summarise(mean_pm10 = mean(pm10, na.rm = TRUE),
            median_pm10 = median(pm10, na.rm = TRUE)) %>%
  right_join(monitor_points)


ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5)  +
  
  # Add road data
  geom_sf(data = roads_filtered, color = "gray", size = 0.3) +  # Add roads data
  
  # Add monitor points with transparency within the bounding box
  geom_sf(data = pm10_summary, aes(geometry = geometry, color = mean_pm10), alpha = 0.7, size = 6) +

  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  
  # Customize
  labs(title = "Mean PM 10 for Each Monitor During Harmattan",
       x = "", 
       y = "") +
  theme_bw() +
  scale_color_gradient(low = "#0540a1", high = "#eb4c2d",
                       breaks = c(min(pm10_summary$mean_pm10, na.rm = TRUE), 210, 365, max(pm10_summary$mean_pm10)),
                       labels = c(paste(round(min(pm10_summary$mean_pm10),0)), "210", "365", paste(round(max(pm10_summary$mean_pm10),0))),
                       name = "PM 10 \n(µg/m³)") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    plot.title = element_text(size = 14), # Increase plot title size
    legend.text = element_text(size = 12), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    legend.key.size = unit(1.3, "lines") # Increase legend tick mark size
  )
```

```{r pm maps NOT during harmattan, echo=FALSE, message=FALSE, warning=FALSE}
# NOT DURING HARMATTTAN ----

pm25_summary <- pm25_corrected %>%
  filter(date < harmattan_start | date >= harmattan_end) %>%
  group_by(monitor) %>%
  summarise(mean_pm25 = mean(pm25, na.rm = TRUE),
            median_pm25 = median(pm25, na.rm = TRUE)) %>%
  right_join(monitor_points)

ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5)  +
  
  # Add road data
  geom_sf(data = roads_filtered, color = "gray", size = 0.3) +  # Add roads data
  
  # Add monitor points with transparency within the bounding box
  geom_sf(data = pm25_summary, aes(geometry = geometry, color = mean_pm25), alpha = 0.7, size = 6) +

  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  
  # Customize
  labs(title = "Mean PM 2.5 for Each Monitor Outside of Harmattan",
       x = "", 
       y = "") +
  theme_bw() +
  scale_color_gradient(low = "#0540a1",  high = "#eb4c2d",
                       breaks = c(min(pm25_summary$mean_pm25, na.rm = TRUE), 45,70, max(pm25_summary$mean_pm25)),
                       labels = c(paste(round(min(pm25_summary$mean_pm25),0)), "45", "70", paste(round(max(pm25_summary$mean_pm25),0))),
                       name = "PM 2.5 \n(µg/m³)") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    plot.title = element_text(size = 14), # Increase plot title size
    legend.text = element_text(size = 12), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    legend.key.size = unit(1.3, "lines") # Increase legend tick mark size
  )

## DECISION TO RECREATE PLOT ABOVE WITH 3 POINT COLOR SCALE SO OUTLIER DOESN'T SKEW VARIABILITY OF OTHER POINTS

ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5)  +
  
  # Add road data
  geom_sf(data = roads_filtered, color = "gray", size = 0.3) +  # Add roads data
  
  # Add monitor points with transparency within the bounding box
  geom_sf(data = pm25_summary, aes(geometry = geometry, color = mean_pm25), alpha = 0.7, size = 6) +

  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  
  # Customize
  labs(title = "Mean PM 2.5 for Each Monitor Outside of Harmattan",
       x = "", 
       y = "") +
  theme_bw() +
  scale_color_gradient2(low = "#0540a1", mid = "#eb4c2d", high = "goldenrod", 
                        midpoint = 55,  # Set the midpoint value
                        breaks = c(min(pm25_summary$mean_pm25, na.rm = TRUE), 55, max(pm25_summary$mean_pm25)),
                        labels = c(paste(round(min(pm25_summary$mean_pm25),0)), "55", paste(round(max(pm25_summary$mean_pm25),0))),
                        name = "PM 2.5 \n(µg/m³)") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    plot.title = element_text(size = 14), # Increase plot title size
    legend.text = element_text(size = 12), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    legend.key.size = unit(1.3, "lines") # Increase legend tick mark size
  )


pm1_summary <- pm1_corrected %>%
  filter(date < harmattan_start | date <= harmattan_end) %>%
  group_by(monitor) %>%
  summarise(mean_pm1 = mean(pm1, na.rm = TRUE),
            median_pm1 = median(pm1, na.rm = TRUE)) %>%
  right_join(monitor_points)


ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5)  +
  
  # Add road data
  geom_sf(data = roads_filtered, color = "gray", size = 0.3) +  # Add roads data
  
  # Add monitor points with transparency within the bounding box
  geom_sf(data = pm1_summary, aes(geometry = geometry, color = mean_pm1), alpha = 0.7, size = 6) +

  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  
  # Customize
  labs(title = "Mean PM 1 for Each Monitor Outside of Harmattan",
       x = "", 
       y = "") +
  theme_bw() +
  scale_color_gradient(low = "#0540a1", high = "#eb4c2d", 
                       breaks = c(min(pm1_summary$mean_pm1), 40, 55, max(pm1_summary$mean_pm1)),
                       labels = c(paste(round(min(pm1_summary$mean_pm1),0)), "40", "55", paste(round(max(pm1_summary$mean_pm1),0))),
                       name = "PM 1 \n(µg/m³)") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    plot.title = element_text(size = 14), # Increase plot title size
    legend.text = element_text(size = 12), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    legend.key.size = unit(1.3, "lines") # Increase legend tick mark size
  )


pm10_summary <- pm10_corrected %>%
  filter(date < harmattan_start | date <= harmattan_end) %>%
  group_by(monitor) %>%
  summarise(mean_pm10 = mean(pm10, na.rm = TRUE),
            median_pm10 = median(pm10, na.rm = TRUE)) %>%
  right_join(monitor_points)


ggplot() +
  # Add Ghana regions within the bounding box
  geom_sf(data = regions, fill = ifelse(regions$region == "Bono East", "#f5e493", "#dadbe0"), color = "black", alpha = 0.5)  +
  
  # Add road data
  geom_sf(data = roads_filtered, color = "gray", size = 0.3) +  # Add roads data
  
  # Add monitor points with transparency within the bounding box
  geom_sf(data = pm10_summary, aes(geometry = geometry, color = mean_pm10), alpha = 0.7, size = 6) +

  # Zoom in to the bounding box
  coord_sf(xlim = c(-2.2, -1.3), ylim = c(7.6, 8.8), expand = FALSE) +
  
  # Customize
  labs(title = "Mean PM 10 for Each Monitor Outside of Harmattan",
       x = "", 
       y = "") +
  theme_bw() +
  scale_color_gradient(low = "#0540a1", high = "#eb4c2d", 
                       breaks = c(min(pm10_summary$mean_pm10, na.rm = TRUE), 140, 230, max(pm10_summary$mean_pm10)),
                       labels = c(paste(round(min(pm10_summary$mean_pm10),0)), "140", "230", paste(round(max(pm10_summary$mean_pm10),0))),
                       name = "PM 10 \n(µg/m³)") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    plot.title = element_text(size = 14), # Increase plot title size
    legend.text = element_text(size = 12), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    legend.key.size = unit(1.3, "lines") # Increase legend tick mark size
  )

```



## ROAD FURTHER EXPLORATION AND CUSTOMIZATION 

```{r prep road data, echo=FALSE, message=FALSE, warning=FALSE}
# groups road types so sample size for regression is better
roads_filtered_grouped <- roads_filtered %>%
  mutate(highway = case_when(highway %in% c("secondary", "trunk", "tertiary") ~ "City Connection",
                             highway == "residential" ~ "Residential",
                             TRUE ~ "Other")) 


# prepare data
missing_monitor_location_sf <- missing_monitor_location %>% select(everything(), geometry) %>% st_as_sf()

# 100 meter buffer: Medgyesi et al. (2023), the 100-meter buffer is relevant to the first-order decay of air pollutants from major roadways and is associated with increased health risks
monitor_buffers <- st_buffer(missing_monitor_location_sf, dist = 100)

roads_within_buffers <- st_intersection(roads_filtered_grouped, monitor_buffers) %>%
  mutate(road_length = st_length(geometry))

#count length of road within 100 m buffer
total_road_length_per_monitor <- roads_within_buffers %>%
  group_by(monitor) %>%
  summarize(total_length = sum(road_length)) %>%
  ungroup() %>%
  st_drop_geometry()

#add road length data to location data 
monitor_with_road_lengths <- left_join(missing_monitor_location_sf, total_road_length_per_monitor, by = "monitor") %>%
  mutate(total_length = as.numeric(total_length)) %>%
  mutate(total_length = case_when(
    is.na(total_length) ~ 0,
    TRUE ~ total_length
  )) 
```

```{r harmattan roads, echo=FALSE, message=FALSE, warning=FALSE}
# calculate monitor summary stats during harmattan to compare seasonality
summary_stats_harmattan <- pm25_corrected %>%
  filter(date >= harmattan_start |
         date < harmattan_end) %>%
    group_by(monitor) %>%
    summarize(
      mean_value = mean(pm25, na.rm = TRUE),
      median_value = median(pm25, na.rm = TRUE),
      sd_value = sd(pm25, na.rm = TRUE)
    )

# Merge with monitor location data
monitor_data <- monitor_with_road_lengths %>%
  left_join(summary_stats_harmattan, by = "monitor")
  
# Convert monitor data to spatial object using sf
monitor_sf <- st_as_sf(monitor_data)
  
# Calculate distance to nearest road and get the nearest road index
nearest_roads <- st_nearest_feature(monitor_sf, roads_filtered_grouped)
distances <- st_distance(monitor_sf, roads_filtered_grouped[nearest_roads, ], by_element = TRUE)
  
# Add distances and nearest road type to the monitor data
monitor_sf <- monitor_sf %>%
  mutate(
    distance_to_road = as.numeric(distances),
    road_type = roads_filtered_grouped$highway[nearest_roads],
    road_surface = roads_filtered_grouped$surface[nearest_roads]

    ) 

close_roads <- monitor_sf %>% 
  mutate(main_road = ifelse(location %in% c("Amoma", "Anyima", "Krabonso", "Ampoma", "Kokuma"), 
                            "Not Main Road", 
                            "Main Road"))



population_road_model <- lm(log(mean_value) ~ log(population) + main_road, data = close_roads)

summary(population_road_model)

# Model summaries
model_name_pop_road <- paste("Log of Mean PM 2.5 Regressed on Population and Primary Road Status")

model_list <- setNames(list(population_road_model), 
                       c(model_name_pop_road))

summary_table <- modelsummary(
  model_list, 
  #coef_rename = c("distance_to_road" = "Distance to Road", "total_length" = "Road Length within 100m Buffer"),
  gof_omit = "AIC|BIC|Log.Lik",
  estimate = "{estimate} ({std.error}){stars}",
  statistic = "p.value"
)

```


```{r population vs pm25 during harmattan, echo=FALSE, message=FALSE, warning=FALSE}
## PLOT OF PM

coefs <- summary(population_road_model)$coefficients
slope <- coefs["log(population)", "Estimate"]
p_value <- coefs["log(population)", "Pr(>|t|)"]
  

# Plot considering road type
close_roads %>%
  #filter(population < 20000) %>%
  ggplot(aes(x = log(population), y = log(mean_value))) +
  geom_point(aes(x = log(population), y = log(mean_value))) +
  geom_smooth(aes(x = log(population), y = log(mean_value)), method = "lm", se = FALSE) +
  annotate("text", x = Inf, y = Inf, label = paste("Slope:", round(slope, 2), "\nP-value:", round(p_value, 3)),
             hjust = 1.7, vjust = 7, size = 4, color = "black") +
  labs(x = "Log of Community Population",
       y = "Log of Average PM 2.5 Reading") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.ticks = element_line(size = 0.7), # Increase tick mark size
    legend.text = element_text(size = 10), # Increase legend text size
    legend.title = element_text(size = 12), # Increase legend title size
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
  )
```


## roads excluding harmattan
```{r non harmattan roads, echo=FALSE, message=FALSE, warning=FALSE}
summary_stats_non_harmattan <- pm25_corrected %>%
  filter(date < harmattan_start |
         date >= harmattan_end) %>%
    group_by(monitor) %>%
    summarize(
      mean_value = mean(pm25, na.rm = TRUE),
      median_value = median(pm25, na.rm = TRUE),
      sd_value = sd(pm25, na.rm = TRUE)
    )

# Merge with monitor location data
monitor_data <- monitor_with_road_lengths %>%
  left_join(summary_stats_non_harmattan, by = "monitor")
  
  # Convert monitor data to spatial object using sf
monitor_sf <- st_as_sf(monitor_data)
  
  # Calculate distance to nearest road and get the nearest road index
nearest_roads <- st_nearest_feature(monitor_sf, roads_filtered_grouped)
distances <- st_distance(monitor_sf, roads_filtered_grouped[nearest_roads, ], by_element = TRUE)
  
  # Add distances and nearest road type to the monitor data
monitor_sf <- monitor_sf %>%
  mutate(
    distance_to_road = as.numeric(distances),
    road_type = roads_filtered_grouped$highway[nearest_roads]
    ) 

close_roads <- monitor_sf %>% 
  #filter(distance_to_road < 200) %>%
  left_join(monitor_community_populations, by = "monitor") %>%
  mutate(main_road = ifelse(location %in% c("Amoma", "Anyima", "Krabonso", "Ampoma", "Kokuma"), 
                            "Not Main Road", 
                            "Main Road"))

population_road_model <-lm(log(mean_value) ~ log(population) + main_road, data = close_roads)

# Model summaries
model_name_pop_road <- paste("Log of Mean PM 2.5 Regressed on Population and Primary Road Status")

model_list <- setNames(list(population_road_model), 
                       c(model_name_pop_road))

summary_table <- modelsummary(
  model_list, 
  coef_rename = c("distance_to_road" = "Distance to Road", "total_length" = "Road Length within 100m Buffer"),
  gof_omit = "AIC|BIC|Log.Lik",
  estimate = "{estimate} ({std.error}){stars}",
  statistic = "p.value"
)
  
  
  

coefs <- summary(population_road_model)$coefficients
slope <- coefs["log(population)", "Estimate"]
p_value <- coefs["log(population)", "Pr(>|t|)"]
  
```

## APPENDIX PLOTS

```{r time series pm25 individual monitor locations facet, echo=FALSE, warning= FALSE, message=FALSE}
monitor_names <- tibble::tribble(
  ~monitor, ~description, ~location,
  "MOD-PM-00887", "Babator", "Babatokuma GH",
  "MOD-PM-00881", "Mo-line", "Kintampo GH",
  "MOD-PM-01052", "Kokuma", "Kokuma GH",
  "MOD-PM-01055", "Krabonso", "Krabonso GH",
  "MOD-PM-00893", "Jato Akura", "Jato Akura GH",
  "MOD-PM-01060", "Kintampo Central College Areas", "Kintampo GH",
  "MOD-PM-00899", "Nante", "Nante GH",
  "MOD-00399", "Dwenewoho", "Dwenewoho GH",
  "MOD-PM-01057", "Sorunuase", "Sorunuase GH",
  "MOD-PM-00877", "Pamdu", "Pamdu GH",
  "MOD-PM-00898", "Anokyekrom", "Anokyekrom GH",
  "MOD-PM-00900", "Kawampe", "Kawampe GH",
  "MOD-PM-00883", "Portor", "Portor GH",
  "MOD-00398", "Kurawura Akura", "Kurawura Akura GH",
  "MOD-PM-00884", "KHRC Residency", "Kintampo GH",
  "MOD-PM-01056", "Gulumpe", "Gulumpe GH",
  "MOD-PM-00879", "Dawadawa", "Dawadawa GH",
  "MOD-PM-00897", "Pramposo", "Pramposo GH",
  "MOD-PM-01053", "Anyima", "Anyima GH",
  "MOD-PM-00880", "Kadelso", "Kadelso GH",
  "MOD-PM-01051", "Weila", "Weila GH",
  "MOD-PM-01058", "Krutakyi", "Krutakyi GH",
  "MOD-PM-00895", "Ntankro", "Kintampo GH",
  "MOD-PM-00889", "Beposo", "Beposo GH",
  "MOD-PM-00882", "Apaaso", "Kintampo GH",
  "MOD-PM-01059", "Paninamisa", "Paninamisa GH",
  "MOD-00397", "Apesika", "Apesika GH",
  "MOD-PM-00878", "Jema Pentecost", "Jema GH",
  "MOD-00401", "New Longoro", "New Longoro GH",
  "MOD-PM-00891", "Habitat", "Kintampo GH",
  "MOD-PM-00885", "Kwabia", "Kwabia GH",
  "MOD-PM-00890", "Akora", "Akora GH",
  "MOD-PM-00876", "Amoma", "Amoma GH",
  "MOD-PM-00888", "Nante Zongo", "Nante Zongo GH",
  "MOD-PM-00894", "Alhassan Akura", "Alhassan Akura GH",
  "MOD-00400", "Kintampo-PTC Area (Magazine)", "Kintampo GH",
  "MOD-PM-01054", "Ampoma", "Ampoma GH",
  "MOD-PM-00896", "Asantekwaa", "Asantekwaa GH",
  "MOD-PM-00886", "Jema Zongo", "Jema GH",
  "MOD-PM-00892", "Atta Akura", "Atta Akura GH"
)

monitor_names <- monitor_names %>%
  mutate(location = gsub(" GH", "", location)) %>%
  mutate(locations = if_else(grepl("Kintampo", location),
                            paste(description, location, sep = ", "),
                            location))




# Custom function to add line breaks at a specific length
add_line_breaks <- function(x, width) {
  sapply(x, function(y) {
    if (nchar(y) > width) {
      paste(strwrap(y, width = width), collapse = "\n")
    } else {
      y
    }
  })
}

# Apply custom function to locations
monitor_names <- monitor_names %>%
  mutate(locations = if_else(grepl("Kintampo", location),
                             paste(description, "Kintampo", sep = "\n"),
                             location)) %>%
  mutate(locations = add_line_breaks(locations, width = 20))

# Plot with updated facet labels
pm25_corrected %>%
  left_join(monitor_names) %>%
  group_by(date, locations) %>%
  summarize(mean_pm25 = mean(pm25, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = mean_pm25)) +
  geom_line(color = "black", alpha = 1) +
  theme_minimal() +
  facet_wrap(~locations) +
  labs(y = "Mean PM 2.5") +
  theme(
    axis.text = element_text(size = 11), # Increase axis text size
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 11)  # Increase facet group text size
  )

```


