---
title: "Build Model to Calibrate Modulairs with GRIMM"
output: 
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
date: "`r Sys.Date()`"
---


```{r load libraries, include=FALSE}
# load libraries 
library(tidyverse)
library(purrr)
library(lubridate)
library(tidyr)
library(sf)
library(here)
library(readxl)
library(performance)
library(lme4)
library(ICC)
library(broom)
library(modelr)
library(viridis)

# ML packages 
library(tidymodels)
library(tictoc)
library(vip)
library(patchwork)
library(doParallel)
library(xgboost)
library(glmnet)
library(mgcv)
library(yardstick) 
library(furrr)
library(future)
```

##### Load and prepare Modulair Data 
```{r, echo=FALSE, message=FALSE, warning=FALSE}

## PM DATA ----
modulair_pm_25_hourly <- read_csv(here("data", "pm", "summarized", "pm25_community_hourly_20241031-20250901.csv")) %>%
  filter(monitor %in% c("MOD-PM-00884")) %>% # filter to monitors co-located with GRIMM
  filter(date > as.Date("2025-08-10")) %>% ## filter data prior to GRIMM deployment 
  filter(!is.na(mean_pm25)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm25))

modulair_pm_1_hourly <- read_csv(here("data", "pm", "summarized", "pm1_community_hourly_20241031-20250901.csv")) %>%
  filter(monitor %in% c("MOD-PM-00884")) %>% # filter to monitors co-located with GRIMM
  filter(date > as.Date("2025-08-10")) %>% ## filter data prior to GRIMM deployment 
  filter(!is.na(mean_pm1)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm1))

modulair_pm_10_hourly <- read_csv(here("data", "pm", "summarized", "pm10_community_hourly_20241031-20250901.csv")) %>%
  filter(monitor %in% c("MOD-PM-00884")) %>% # filter to monitors co-located with GRIMM
  filter(date > as.Date("2025-08-10")) %>% ## filter data prior to GRIMM deployment 
  filter(!is.na(mean_pm10)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm10))

batch1_pm_00884_hourly <- left_join(modulair_pm_25_hourly, modulair_pm_1_hourly, by = c("monitor", "date", "hour")) %>%
  left_join(modulair_pm_10_hourly, by = c("monitor", "date", "hour"))

# 
# batch1_pm_00884_hourly <- read_csv(here("data", "pm", "summarized", "modpm00884_raw_hourly_20250801-20250901.csv")) %>%
#   rename(mean_pm1 = mean_pm1_raw,
#          mean_pm10 = mean_pm10_raw,
#          mean_pm25 = mean_pm25_raw) %>%
#   filter(!is.na(mean_pm25))


batch2_pm_00884_hourly <- read_csv(here("data", "pm", "summarized", "modpm00884_hourly_20250901-20251231.csv")) %>%
  filter(!is.na(mean_pm25))




monitor_start_dates <- tibble::tibble(
  monitor = c("MOD-PM-00884", "MOD-00077", "MOD-00397", "MOD-PM-00879", "MOD-PM-00893"),
  start_date = as.Date(c("2025-08-10", "2025-12-15", "2026-01-20", "2026-02-13", "2026-02-13"))
)

batch3_modulair_pm_25_hourly <- read_csv(here("data", "pm", "summarized", "pm25_grimm_colocated_hourly_20251215-20260301.csv")) %>%
  filter(!is.na(mean_pm25)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm25)) %>%
  left_join(monitor_start_dates, by = "monitor") %>%
  filter(date >= start_date | is.na(start_date)) %>%
  select(-start_date)
  

batch3_modulair_pm_1_hourly <- read_csv(here("data", "pm", "summarized", "pm1_grimm_colocated_hourly_20251215-20260301.csv")) %>%
  filter(!is.na(mean_pm1)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm1)) %>% 
  left_join(monitor_start_dates, by = "monitor") %>%
  filter(date >= start_date | is.na(start_date)) %>%
  select(-start_date)
  

batch3_modulair_pm_10_hourly <- read_csv(here("data", "pm", "summarized", "pm10_grimm_colocated_hourly_20251215-20260301.csv")) %>%
  filter(!is.na(mean_pm10)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm10)) %>% 
  left_join(monitor_start_dates, by = "monitor") %>%
  filter(date >= start_date | is.na(start_date)) %>%
  select(-start_date)


batch3_hourly <- left_join(batch3_modulair_pm_25_hourly, batch3_modulair_pm_1_hourly, by = c("monitor", "date", "hour")) %>%
  left_join(batch3_modulair_pm_10_hourly, by = c("monitor", "date", "hour"))


modulair_pm_hourly <- bind_rows(batch1_pm_00884_hourly, batch2_pm_00884_hourly, batch3_hourly) %>%
  distinct(date, hour, monitor, .keep_all = TRUE) # remove overlapping day of sept 1




## WEATHER DATA -----

modulair_temp_hourly <- read_csv(here("data", "weather", "summarized", "temp_hourly_20241031-20250901.csv")) %>%
  filter(monitor %in% c("MOD-PM-00884")) %>% # filter to monitors co-located with GRIMM
  filter(date > as.Date("2025-08-10")) %>% ## filter data prior to GRIMM deployment 
  filter(!is.na(mean_met_temp))%>%
  select(-c(n_minute_obs, n_active, fleet_average_met_temp)) 

modulair_rh_hourly <- read_csv(here("data", "weather", "summarized", "rh_hourly_20241031-20250901.csv")) %>%
  filter(monitor %in% c("MOD-PM-00884")) %>% # filter to monitors co-located with GRIMM
  filter(date > as.Date("2025-08-10")) %>% ## filter data prior to GRIMM deployment 
  filter(!is.na(mean_met_rh))%>%
  select(-c(n_minute_obs, n_active, fleet_average_met_rh))

batch1_weather_00884_hourly <- left_join(modulair_temp_hourly, modulair_rh_hourly, by = c("monitor", "date", "hour"))

batch2_weather_00884_hourly <- read_csv(here("data", "weather", "summarized", "modpm00884_hourly_20250901-20251231.csv"))


batch3_weather_hourly <- read_csv(here("data", "weather", "summarized", "grimm_weather_hourly_20251215-20260301.csv")) %>% 
  left_join(monitor_start_dates, by = "monitor") %>%
  filter(date >= start_date | is.na(start_date)) %>%
  select(-start_date)



modulair_weather_hourly <- bind_rows(batch1_weather_00884_hourly, batch2_weather_00884_hourly, batch3_weather_hourly) %>%
  distinct(date, hour, monitor, .keep_all = TRUE) %>% # remove overlapping day of sept 1
  filter(!is.na(mean_met_temp) & !is.na(mean_met_rh))




## FULL DATA (PM AND WEATHER) ----

modulair_full <- full_join(modulair_pm_hourly, modulair_weather_hourly, by = c("monitor", "date", "hour")) %>%
  mutate(
    date_hour = as.POSIXct(paste(date, hour), format = "%Y-%m-%d %H", tz = "UTC") # create date hour column for GRIMM modeling
  ) %>%
  select(monitor, date_hour, date, hour, everything()) %>%
  rename(mod_pm25 = mean_pm25,
         mod_pm1 = mean_pm1,
         mod_pm10 = mean_pm10,
         mod_temp = mean_met_temp,
         mod_rh = mean_met_rh) %>%
  filter(date < as.Date("2026-03-02"))
```


##### Load in the GRIMM data 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
grimm_1 <- read_csv(here("data", "FEMs", "GRIMM", "clean", "grimm_hourly_20250811-20251008.csv")) %>%
  mutate(hour = hour(date_hour)) %>%
  select(date_hour, date, hour, everything())

grimm_2 <- read_csv(here("data", "FEMs", "GRIMM", "clean", "grimm_hourly_20251008-20251218.csv")) %>%
  mutate(hour = hour(date_hour)) %>%
  select(date_hour, date, hour, everything())

grimm_3 <- read_csv(here("data", "FEMs", "GRIMM", "clean", "grimm_hourly_20260123_20260206.csv")) %>%
  mutate(hour = hour(date_hour)) %>%
  select(date_hour, date, hour, everything())

grimm_full <- bind_rows(grimm_1, grimm_2, grimm_3)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}

modulair_full %>% 
  filter(mod_pm10 < 250) %>%
  ggplot(aes(x = mod_pm10)) +
  geom_histogram(bins = 20, color = "white") +
  theme_minimal() +
  labs(title = "Histogram of Modulair PM 10",
       x = "Modulair PM 10 (filtered below 500 to remove outliers)")

grimm_full %>%
  ggplot(aes(x = grimm_pm10)) +
  geom_histogram(bins = 20, color = "white") +
  theme_minimal() +
  labs(title = "Histogram of GRIMM PM 10",
       x = "GRIMM PM 10",
       y = "Count") +
  coord_cartesian(xlim = c(0, 250))
```

##### Merge modular and GRIMM data 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# merge and add in seasonality for harmattan
mod_grimm_full <- inner_join(modulair_full, grimm_full, by = c("date_hour", "date", "hour")) %>%
  mutate(month = month(date),
         day = day(date),
         season = case_when(month %in% c(12, 1, 2) ~ "Harmattan",
                            month %in% c(3, 4, 5, 6, 7, 8, 9, 10, 11) ~ "Not Harmattan")) %>%
  mutate(season = as.factor(season))
  
```


## Comparison of GRIMM to Modulair reference 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
metrics_df <- mod_grimm_full %>%
  #filter(monitor != "MOD-PM-00884") %>%
  yardstick::metrics(truth = grimm_pm10, estimate = mod_pm10)

rmse_val <- metrics_df %>% filter(.metric == "rmse") %>% pull(.estimate)
rsq_val  <- metrics_df %>% filter(.metric == "rsq") %>% pull(.estimate)

slope_val <- mod_grimm_full %>%
  # filter(monitor != "MOD-PM-00884") %>%
  lm(mod_pm10 ~ grimm_pm10, data = .) %>%
  coef() %>%
  .[2]


mod_grimm_full %>%
  # filter(monitor != "MOD-PM-00884") %>%
  ggplot(aes(x = grimm_pm10, y = mod_pm10)) +
  geom_point(alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  geom_smooth(method = "lm", se = FALSE, color = "black", linetype = "dashed") +
  annotate(
  "label",
  x = Inf, y = -Inf,
  hjust = 1.1, vjust = -0.5,
  label = paste0(
    "RMSE: ", round(rmse_val, 2),
    "\nPearson R²: ", round(rsq_val, 2),
    "\nSlope: ", round(slope_val, 2)
  ),
  size = 4,
  label.size = .1
) +
  theme_minimal() +
  labs(
    x = "GRIMM PM 10",
    y = "Modulair colocated PM 10",
    title = "Baseline comparison of GRIMM + reference Modulair PM 10",
    subtitle = "Modulair data was corrected based on fleet average colocation \nRed line shows slope 1"
  )



## BLAND ALTMAN PLOT
ba_raw <- mod_grimm_full %>%
  mutate(
    mean_pair = (mod_pm10 + grimm_pm10) / 2,
    diff_pair = mod_pm10 - grimm_pm10
  )

mean_diff <- mean(ba_raw$diff_pair, na.rm = TRUE)
sd_diff   <- sd(ba_raw$diff_pair, na.rm = TRUE)

ggplot(ba_raw, aes(x = mean_pair, y = diff_pair)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = mean_diff, linetype = "solid") +
  geom_hline(yintercept = mean_diff + 1.96 * sd_diff, linetype = "dashed") +
  geom_hline(yintercept = mean_diff - 1.96 * sd_diff, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Bland–Altman Plot: Raw Modulair vs GRIMM",
    x = "Mean of Modulair and GRIMM PM10",
    y = "Difference (Modulair - GRIMM)"
  )
```


## Set up day blocked train/test split for ML

Random sample of days, with all hours from a selected day assigned entirely to either the training or test set. This removes leakage from similar times, and hours with multiple observations (e.g. from two monitors that recorded at the same time) can't belong in both test and train.  

**Note on time-based validation.**
Another option would be to split the data by time (e.g. training the model on earlier dates and testing it on a later continuous block of time). Since weather conditions and pollution events tend to change gradually, observations that are close together in time can be very similar, so if nearby hours end up in both the training and test sets, it can lead to better performance because the model is seeing very similar conditions during training. 

I didn’t implement that approach here because we don't have much colocation data across both seasons, and monitors were added at different times.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(123)

test_dates <- mod_grimm_full %>%
  distinct(date, season) %>%
  group_by(season) %>%
  slice_sample(prop = 0.2) %>%
  ungroup() %>%
  pull(date)

train_day <- mod_grimm_full %>% filter(!date %in% test_dates)
test_day  <- mod_grimm_full %>% filter(date %in% test_dates)

cv_folds_day <- group_vfold_cv(train_day, group = date, v = 5, strata = season)
```


## Elastic Net Regression

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=10}
calib_recipe <- recipe(
  grimm_pm10 ~ mod_pm10 + mod_temp + mod_rh + hour + season,
  data = train_day
) %>%
  step_mutate(
    hour_num = as.integer(hour),
    sin_hour = sin(2*pi*hour_num/24),
    cos_hour = cos(2*pi*hour_num/24)
  ) %>%
  step_rm(hour, hour_num) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_impute_bag(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_zv(all_predictors())

enet_spec <- linear_reg(
  penalty = tune(),     # overall shrinkage
  mixture = tune()      # 0 = ridge, 1 = LASSO, between = elastic net
) %>%
  set_engine("glmnet")

enet_grid <- grid_regular(
  penalty(range = c(-4, 0)),   # 10^-4 to 10^0 = 1e-4 to 1.0
  mixture(range = c(0, 1)),
  levels = 6
)

enet_wf <- workflow() %>%
  add_recipe(calib_recipe) %>%
  add_model(enet_spec)

set.seed(123)
enet_tuned <- tune_grid(
  enet_wf,
  resamples = cv_folds_day,
  grid = enet_grid,
  metrics = metric_set(yardstick::rmse, yardstick::rsq, yardstick::mae)
)

best_enet <- select_best(enet_tuned, metric = "rmse")
final_enet_wf <- finalize_workflow(enet_wf, best_enet)

# Fit on training data
enet_fit_train <- fit(final_enet_wf, data = train_day)

# Predict on test data
enet_preds <- predict(enet_fit_train, new_data = test_day) %>%
  bind_cols(test_day %>% select(grimm_pm10))

metrics_df <- enet_preds %>%
  yardstick::metrics(truth = grimm_pm10, estimate = .pred)

rmse_val <- metrics_df %>% filter(.metric == "rmse") %>% pull(.estimate)
mae_val  <- metrics_df %>% filter(.metric == "mae")  %>% pull(.estimate)

pearson_r2 <- cor(enet_preds$grimm_pm10, enet_preds$.pred, use = "complete.obs")^2

test_r2 <- 1 - sum((enet_preds$grimm_pm10 - enet_preds$.pred)^2) /
  sum((enet_preds$grimm_pm10 - mean(enet_preds$grimm_pm10))^2)

# calibration
cal_fit   <- lm(grimm_pm10 ~ .pred, data = enet_preds)
cal_int   <- coef(cal_fit)[1]
cal_slope <- coef(cal_fit)[2]

# shared axis limits
lims <- range(c(enet_preds$grimm_pm10, enet_preds$.pred), na.rm = TRUE)

caption_text <- paste0(
  "Test-set performance: RMSE = ", round(rmse_val, 2),
  "\nMAE = ", round(mae_val, 2),
  "\nPearson R² = ", round(pearson_r2, 2),
  "\nTest-set R² = ", round(test_r2, 2),
  "\nCalibration slope (Obs~Pred) = ", round(cal_slope, 2),
  "\nIntercept = ", round(cal_int, 2), "."
)

ggplot(enet_preds, aes(x = .pred, y = grimm_pm10)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "black") +
  coord_equal(xlim = lims, ylim = lims) +
  labs(
    title = "Elastic Net: Predicted vs Observed PM10 (Day-blocked Test Set)",
    x = "Predicted PM10 (Elastic Net)",
    y = "Observed PM10 (GRIMM)",
    subtitle = "Red line = 1:1 identity; dashed line = regression fit",
    caption = caption_text
  ) +
  theme_minimal()


# BLAND ALTMAN ---
ba_enet <- enet_preds %>%
  mutate(
    mean_pair = (.pred + grimm_pm10) / 2,
    diff_pair = .pred - grimm_pm10
  )

mean_diff <- mean(ba_enet$diff_pair, na.rm = TRUE)
sd_diff   <- sd(ba_enet$diff_pair, na.rm = TRUE)

ggplot(ba_enet, aes(x = mean_pair, y = diff_pair)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = mean_diff, linetype = "solid") +
  geom_hline(yintercept = mean_diff + 1.96 * sd_diff, linetype = "dashed") +
  geom_hline(yintercept = mean_diff - 1.96 * sd_diff, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Bland–Altman Plot: Elastic Net Predictions vs GRIMM",
    x = "Mean of Predicted and Observed PM10",
    y = "Difference (Predicted - Observed)"
  )

## NORMALIZED RMSE
rmse_val <- rmse_vec(enet_preds$grimm_pm10, enet_preds$.pred)
mean_obs <- mean(enet_preds$grimm_pm10, na.rm = TRUE)

nrmse_pct <- 100 * rmse_val / mean_obs
nrmse_pct


### SAVE MODEL
enet_fit_full <- fit(final_enet_wf, data = mod_grimm_full)

enet_fit_full %>%
  extract_fit_parsnip() %>%
  tidy()

saveRDS(enet_fit_full, here("data", "calibration", "modulair_grimm_enet_model.rds"))

```


## XG Boost 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
xgb_recipe <- recipe(grimm_pm10 ~ mod_pm10 + mod_temp + mod_rh + hour + season, data = train_day) %>%
  step_mutate(
    season   = factor(season),
    hour_num = as.integer(hour),
    sin_hour = sin(2*pi*hour_num/24),
    cos_hour = cos(2*pi*hour_num/24)
  ) %>%
  step_rm(hour, hour_num) %>%
  step_novel(season) %>%   #  for future prediction if unseen season in new data
  step_dummy(season, one_hot = TRUE) %>%
  step_impute_bag(all_numeric_predictors()) %>% # fix for missingness. Imputes using bagged decision trees. 
  step_zv(all_predictors()) # drop zero-variance (e.g. season when all non-harmattan)

xgb_spec <- boost_tree(
  trees = 1000,
  tree_depth = tune(),
  learn_rate = tune(),
  loss_reduction = tune(),
  sample_size = tune(),
  mtry = tune()
  ) %>%
  set_mode("regression") %>%
  set_engine("xgboost")

xgb_grid <- grid_max_entropy(
  tree_depth(),
  learn_rate(),
  loss_reduction(),
  sample_prop(),
  finalize(mtry(), train_day),
  size = 25
)

xgb_wf <- workflow() %>%
  add_recipe(xgb_recipe) %>%
  add_model(xgb_spec)


registerDoParallel()
set.seed(123)

xgb_tuned <- tune_grid(
  xgb_wf,
  resamples = cv_folds_day,
  grid = xgb_grid,
  metrics = metric_set(yardstick::rmse, yardstick::rsq, yardstick::mae)
)

best_params <- select_best(xgb_tuned, metric = "rmse")

final_wf <- finalize_workflow(xgb_wf, best_params)

# Fit on TRAIN, evaluate on TEST (custom split => no last_fit)
xgb_fit_train <- fit(final_wf, data = train_day)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=10}
xgb_preds <- predict(xgb_fit_train, new_data = test_day) %>%
  bind_cols(test_day %>% select(grimm_pm10))

metrics_df <- xgb_preds %>%
  yardstick::metrics(truth = grimm_pm10, estimate = .pred)

rmse_val <- metrics_df %>% filter(.metric == "rmse") %>% pull(.estimate)
mae_val  <- metrics_df %>% filter(.metric == "mae")  %>% pull(.estimate)

pearson_r2 <- cor(xgb_preds$grimm_pm10, xgb_preds$.pred, use = "complete.obs")^2

test_r2 <- 1 - sum((xgb_preds$grimm_pm10 - xgb_preds$.pred)^2) /
  sum((xgb_preds$grimm_pm10 - mean(xgb_preds$grimm_pm10))^2)


# calibration
cal_fit   <- lm(grimm_pm10 ~ .pred, data = xgb_preds)
cal_int   <- coef(cal_fit)[1]
cal_slope <- coef(cal_fit)[2]

# shared axis limits
lims <- range(c(xgb_preds$grimm_pm10, xgb_preds$.pred), na.rm = TRUE)

caption_text <- paste0(
  "Test-set performance: RMSE = ", round(rmse_val, 2),
  "\nMAE = ", round(mae_val, 2),
  "\nPearson R² = ", round(pearson_r2, 2),
  "\nTest-set R² = ", round(test_r2, 2),
  "\nCalibration slope (Obs~Pred) = ", round(cal_slope, 2),
  "\nIntercept = ", round(cal_int, 2), "."
)

ggplot(xgb_preds, aes(x = .pred, y = grimm_pm10)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "black") +
  coord_equal(xlim = lims, ylim = lims) +
  labs(
    title = "XGBoost: Predicted vs Observed PM10 (Day-blocked Test Set)",
    x = "Predicted PM10 (XGBoost)",
    y = "Observed PM10 (GRIMM)",
    subtitle = "Red line = 1:1 identity; dashed line = regression fit",
    caption = caption_text
  ) +
  theme_minimal()



## BLAND ALTMAN ---- 

ba_xgb <- xgb_preds %>%
  mutate(
    mean_pair = (.pred + grimm_pm10) / 2,
    diff_pair = .pred - grimm_pm10
  )

mean_diff <- mean(ba_xgb$diff_pair, na.rm = TRUE)
sd_diff   <- sd(ba_xgb$diff_pair, na.rm = TRUE)

ggplot(ba_xgb, aes(x = mean_pair, y = diff_pair)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = mean_diff, linetype = "solid") +
  geom_hline(yintercept = mean_diff + 1.96 * sd_diff, linetype = "dashed") +
  geom_hline(yintercept = mean_diff - 1.96 * sd_diff, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Bland–Altman Plot: XG Boost Predictions vs GRIMM",
    x = "Mean of Predicted and Observed PM10",
    y = "Difference (Predicted - Observed)"
  )



## Normalized RMSE -----
rmse_val <- rmse_vec(xgb_preds$grimm_pm10, xgb_preds$.pred)
mean_obs <- mean(xgb_preds$grimm_pm10, na.rm = TRUE)

nrmse_pct <- 100 * rmse_val / mean_obs
nrmse_pct
```

```{r}
## SAVE XG BOOST MODEL

# Refit final workflow on the full calibration data set
xgb_fit_full <- fit(final_wf, data = mod_grimm_full)

# Save it to disk
saveRDS(xgb_fit_full, here("data", "calibration", "modulair_grimm_xgb_calib_workflow.rds"))
```


#### Save Test/Train splits to run GMR in Python

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## Prep GMR data to run in Python
model_export <- mod_grimm_full %>%
  mutate(
    row_id = row_number(),
    hour_num = as.integer(hour),
    sin_hour = sin(2*pi*hour_num/24),
    cos_hour = cos(2*pi*hour_num/24),
    season_binary = if_else(season == "Harmattan", 1, 0)
  ) %>%
  select(
    row_id, monitor, date_hour, date, hour, season,
    sin_hour, cos_hour, season_binary,
    mod_pm25, mod_pm1, mod_pm10, mod_temp, mod_rh,
    grimm_pm10
  ) %>%
  filter(!is.na(mod_pm10), !is.na(grimm_pm10))


# Day-blocked split export
model_export_day_block_split <- model_export %>%
  mutate(
    split_day_block = case_when(
      date %in% test_dates  ~ "test",
      !date %in% test_dates ~ "train"
    )
  )

readr::write_csv(
  model_export_day_block_split,
  here("data", "calibration", "mod_grimm_day_block_split_pm10_20260315.csv")
)

```
