---
title: "Calibrate Modulairs"
output: 
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
date: "`r Sys.Date()`"
---


This report applies the saved GRIMM calibration models to the broader Modulair hourly data set and summarizes how the calibrated predictions behave. The goals are to compare raw and calibrated values, flag observations that fall outside the training domain, and inspect whether the size of the correction varies by monitor or environmental conditions.

```{r load libraries, include=FALSE}
# load libraries 
library(tidyverse)
library(purrr)
library(lubridate)
library(here)
library(viridis)
library(tidymodels)
library(yardstick) 
library(DT)
```

### Reconstruct calibration training data

This section rebuilds the co-located Modulair-GRIMM training data used to fit the calibration models. These data are used here only to define the range of predictor values seen during training.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## PM DATA ----
modulair_pm_25_hourly <- read_csv(here("data", "pm", "summarized", "pm25_community_hourly_20240816-20250901.csv")) %>%
  filter(monitor %in% c("MOD-PM-00884")) %>% # filter to monitors co-located with GRIMM
  filter(date > as.Date("2025-08-10")) %>% ## filter data prior to GRIMM deployment 
  filter(!is.na(mean_pm25)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm25))

modulair_pm_1_hourly <- read_csv(here("data", "pm", "summarized", "pm1_community_hourly_20240816-20250901.csv")) %>%
  filter(monitor %in% c("MOD-PM-00884")) %>% # filter to monitors co-located with GRIMM
  filter(date > as.Date("2025-08-10")) %>% ## filter data prior to GRIMM deployment 
  filter(!is.na(mean_pm1)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm1))

modulair_pm_10_hourly <- read_csv(here("data", "pm", "summarized", "pm10_community_hourly_20240816-20250901.csv")) %>%
  filter(monitor %in% c("MOD-PM-00884")) %>% # filter to monitors co-located with GRIMM
  filter(date > as.Date("2025-08-10")) %>% ## filter data prior to GRIMM deployment 
  filter(!is.na(mean_pm10)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm10))

batch1_pm_00884_hourly <- left_join(modulair_pm_25_hourly, modulair_pm_1_hourly, by = c("monitor", "date", "hour")) %>%
  left_join(modulair_pm_10_hourly, by = c("monitor", "date", "hour"))

batch2_pm_00884_hourly <- read_csv(here("data", "pm", "summarized", "modpm00884_hourly_20250901-20251231.csv")) %>%
  filter(!is.na(mean_pm25))

monitor_start_dates <- tibble::tibble(
  monitor = c("MOD-PM-00884", "MOD-00077", "MOD-00397", "MOD-PM-00879", "MOD-PM-00893"),
  start_date = as.Date(c("2025-08-10", "2025-12-15", "2026-01-20", "2026-02-13", "2026-02-13"))
)

batch3_modulair_pm_25_hourly <- read_csv(here("data", "pm", "summarized", "pm25_grimm_colocated_hourly_20251215-20260301.csv")) %>%
  filter(!is.na(mean_pm25)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm25)) %>%
  left_join(monitor_start_dates, by = "monitor") %>%
  filter(date >= start_date | is.na(start_date)) %>%
  select(-start_date)

batch3_modulair_pm_1_hourly <- read_csv(here("data", "pm", "summarized", "pm1_grimm_colocated_hourly_20251215-20260301.csv")) %>%
  filter(!is.na(mean_pm1)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm1)) %>% 
  left_join(monitor_start_dates, by = "monitor") %>%
  filter(date >= start_date | is.na(start_date)) %>%
  select(-start_date)

batch3_modulair_pm_10_hourly <- read_csv(here("data", "pm", "summarized", "pm10_grimm_colocated_hourly_20251215-20260301.csv")) %>%
  filter(!is.na(mean_pm10)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm10)) %>% 
  left_join(monitor_start_dates, by = "monitor") %>%
  filter(date >= start_date | is.na(start_date)) %>%
  select(-start_date)

batch3_hourly <- left_join(batch3_modulair_pm_25_hourly, batch3_modulair_pm_1_hourly, by = c("monitor", "date", "hour")) %>%
  left_join(batch3_modulair_pm_10_hourly, by = c("monitor", "date", "hour"))

modulair_pm_hourly <- bind_rows(batch1_pm_00884_hourly, batch2_pm_00884_hourly, batch3_hourly) %>%
  distinct(date, hour, monitor, .keep_all = TRUE) # remove overlapping day of sept 1

## WEATHER DATA -----

modulair_temp_hourly <- read_csv(here("data", "weather", "summarized", "temp_hourly_20240816-20250901.csv")) %>%
  filter(monitor %in% c("MOD-PM-00884")) %>% # filter to monitors co-located with GRIMM
  filter(date > as.Date("2025-08-10")) %>% ## filter data prior to GRIMM deployment 
  filter(!is.na(mean_met_temp))%>%
  select(-c(n_minute_obs, n_active, fleet_average_met_temp)) 

modulair_rh_hourly <- read_csv(here("data", "weather", "summarized", "rh_hourly_20240816-20250901.csv")) %>%
  filter(monitor %in% c("MOD-PM-00884")) %>% # filter to monitors co-located with GRIMM
  filter(date > as.Date("2025-08-10")) %>% ## filter data prior to GRIMM deployment 
  filter(!is.na(mean_met_rh))%>%
  select(-c(n_minute_obs, n_active, fleet_average_met_rh))

batch1_weather_00884_hourly <- left_join(modulair_temp_hourly, modulair_rh_hourly, by = c("monitor", "date", "hour"))

batch2_weather_00884_hourly <- read_csv(here("data", "weather", "summarized", "modpm00884_hourly_20250901-20251231.csv"))

batch3_weather_hourly <- read_csv(here("data", "weather", "summarized", "grimm_weather_hourly_20251215-20260301.csv")) %>% 
  left_join(monitor_start_dates, by = "monitor") %>%
  filter(date >= start_date | is.na(start_date)) %>%
  select(-start_date)

modulair_weather_hourly <- bind_rows(batch1_weather_00884_hourly, batch2_weather_00884_hourly, batch3_weather_hourly) %>%
  distinct(date, hour, monitor, .keep_all = TRUE) %>% # remove overlapping day of sept 1
  filter(!is.na(mean_met_temp) & !is.na(mean_met_rh))

## FULL DATA (PM AND WEATHER) ----

modulair_full <- full_join(modulair_pm_hourly, modulair_weather_hourly, by = c("monitor", "date", "hour")) %>%
  mutate(
    date_hour = as.POSIXct(paste(date, hour), format = "%Y-%m-%d %H", tz = "UTC") # create date hour column for GRIMM modeling
  ) %>%
  select(monitor, date_hour, date, hour, everything()) %>%
  rename(mod_pm25 = mean_pm25,
         mod_pm1 = mean_pm1,
         mod_pm10 = mean_pm10,
         mod_temp = mean_met_temp,
         mod_rh = mean_met_rh) %>%
  filter(date < as.Date("2026-03-02"))

##### Load in the GRIMM data 

grimm_1 <- read_csv(here("data", "FEMs", "GRIMM", "clean", "grimm_hourly_20250811-20251008.csv")) %>%
  mutate(hour = hour(date_hour)) %>%
  select(date_hour, date, hour, everything())

grimm_2 <- read_csv(here("data", "FEMs", "GRIMM", "clean", "grimm_hourly_20251008-20251218.csv")) %>%
  mutate(hour = hour(date_hour)) %>%
  select(date_hour, date, hour, everything())

grimm_3 <- read_csv(here("data", "FEMs", "GRIMM", "clean", "grimm_hourly_20260123_20260206.csv")) %>%
  mutate(hour = hour(date_hour)) %>%
  select(date_hour, date, hour, everything())

grimm_full <- bind_rows(grimm_1, grimm_2, grimm_3)

##### Merge modular and GRIMM data 

# merge and add in seasonality for harmattan
mod_grimm_full <- inner_join(modulair_full, grimm_full, by = c("date_hour", "date", "hour")) %>%
  mutate(month = month(date),
         day = day(date),
         season = case_when(month %in% c(12, 1, 2) ~ "Harmattan",
                            month %in% c(3, 4, 5, 6, 7, 8, 9, 10, 11) ~ "Not Harmattan")) %>%
  mutate(season = as.factor(season))
  
```

## Load and prepare Modulair Data 

Load the broader Modulair data set to which the saved calibration models will be applied.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
modulair_pm_25_hourly_1 <- read_csv(here("data", "pm", "summarized", "pm25_community_hourly_20231024-20240816.csv")) %>%
  filter(!is.na(mean_pm25)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm25))

modulair_pm_25_hourly_2 <- read_csv(here("data", "pm", "summarized", "pm25_community_hourly_20240816-20250901.csv")) %>%
  filter(!is.na(mean_pm25)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_pm25))

modulair_pm_25_hourly <- bind_rows(modulair_pm_25_hourly_1, modulair_pm_25_hourly_2)

modulair_temp_hourly_1 <- read_csv(here("data", "weather", "summarized", "temp_hourly_20230815-20240820.csv")) %>%
  filter(!is.na(mean_met_temp)) %>%
  select(-c(n_minute_obs, fleet_average_met_temp))

modulair_temp_hourly_2 <- read_csv(here("data", "weather", "summarized", "temp_hourly_20240816-20250901.csv")) %>%
  filter(!is.na(mean_met_temp)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_met_temp))

modulair_temp_hourly <- bind_rows(modulair_temp_hourly_1, modulair_temp_hourly_2)


modulair_rh_hourly_1 <- read_csv(here("data", "weather", "summarized", "rh_hourly_20230815-20240820.csv")) %>%
  filter(!is.na(mean_met_rh)) %>%
  select(-c(n_minute_obs, fleet_average_met_rh))

modulair_rh_hourly_2 <- read_csv(here("data", "weather", "summarized", "rh_hourly_20240816-20250901.csv")) %>%
  filter(!is.na(mean_met_rh)) %>%
  select(-c(n_minute_obs, n_active, fleet_average_met_rh))

modulair_rh_hourly <- bind_rows(modulair_rh_hourly_1, modulair_rh_hourly_2)
```

## Join individual variables to full data and format data to match calibration model expectations 

Combines PM2.5, temperature, and relative humidity into one hourly data set.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
modulair_full <- full_join(modulair_pm_25_hourly, modulair_temp_hourly, by = c("monitor", "date", "hour")) %>%
  full_join(modulair_rh_hourly, by = c("monitor", "date", "hour")) %>%
  mutate(
    date_hour = as.POSIXct(paste(date, hour), format = "%Y-%m-%d %H", tz = "UTC")
  ) %>%
  select(monitor, date_hour, date, hour, everything()) %>%
  rename(
    mod_pm25 = mean_pm25,
    mod_temp = mean_met_temp,
    mod_rh   = mean_met_rh
  ) %>%
  mutate(
    month = month(date),
    day   = day(date),
    season = case_when(
      month %in% c(12, 1, 2) ~ "Harmattan",
      TRUE ~ "Not Harmattan"
    ),
    season = factor(season, levels = c("Harmattan", "Not Harmattan"))
  ) %>%
  filter(!is.na(mod_pm25))
```


## Load saved calibration models

These are the previously trained Elastic Net and XGBoost workflows from build_GRIMM_calibration_model.Rmd

```{r, echo=FALSE, message=FALSE, warning=FALSE}
xgb_model <- read_rds(here("data", "calibration", "modulair_grimm_xgb_calib_workflow.rds"))

enet_model <- read_rds(here("data", "calibration", "modulair_grimm_enet_model.rds"))

```


## Apply calibration models

Both saved models are applied to the full Modulair data set. Predicted calibrated PM2.5 values are constrained to be non-negative.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
modulair_preds <- modulair_full %>%
  bind_cols(
    predict(enet_model, new_data = .) %>% rename(pm25_fem_enet = .pred),
    predict(xgb_model, new_data = .) %>% rename(pm25_fem_xgb = .pred)
  ) %>%
  mutate(
    pm25_fem_enet = pmax(pm25_fem_enet, 0), 
    pm25_fem_xgb  = pmax(pm25_fem_xgb, 0)
  )
```


## Compare raw and calibrated PM2.5

### Elastic Net predictions

```{r, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(123)
plot_df <- modulair_preds %>% 
  sample_n(10000)   # or sample_frac(0.05)

## ELASTIC NET PLOT -----
ggplot(plot_df, aes(x = mod_pm25, y = pm25_fem_enet, color = season)) +
  geom_point(alpha = 0.3, size = 1) +
  theme_minimal() +
  labs(x = "Modulair PM 2.5 readings",
       y = "GRIMM calibrated predictions",
       title = "Elastic Net GRIMM calibrated predictions of PM")
```

## XG Boost predictions
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(plot_df, aes(x = mod_pm25, y = pm25_fem_xgb, color = season)) +
  geom_point(alpha = 0.3, size = 1) +
  theme_minimal() +
  labs(x = "Modulair PM 2.5 readings",
       y = "GRIMM calibrated predictions",
       title = "XG Boost GRIMM calibrated predictions of PM")
```



## Flag observations outside the training domain

This section compares the full application data to the predictor ranges observed in the co-located training data. Observations outside those ranges are flagged as extrapolations.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
train_ranges <- mod_grimm_full %>%
  summarise(
    train_pm25_min = min(mod_pm25, na.rm = TRUE),
    train_pm25_max = max(mod_pm25, na.rm = TRUE),
    train_temp_min = min(mod_temp, na.rm = TRUE),
    train_temp_max = max(mod_temp, na.rm = TRUE),
    train_rh_min   = min(mod_rh, na.rm = TRUE),
    train_rh_max   = max(mod_rh, na.rm = TRUE)
  )


train_pm25_min <- train_ranges$train_pm25_min
train_pm25_max <- train_ranges$train_pm25_max
train_temp_min <- train_ranges$train_temp_min
train_temp_max <- train_ranges$train_temp_max
train_rh_min   <- train_ranges$train_rh_min
train_rh_max   <- train_ranges$train_rh_max

modulair_preds_flagged <- modulair_preds %>%
  mutate(
    out_pm25 = mod_pm25 < train_pm25_min | mod_pm25 > train_pm25_max,
    out_temp = !is.na(mod_temp) & (mod_temp < train_temp_min | mod_temp > train_temp_max),
    out_rh   = !is.na(mod_rh)   & (mod_rh   < train_rh_min   | mod_rh   > train_rh_max),
    any_extrapolation = out_pm25 | out_temp | out_rh
  )

overall_extrap_summary <- modulair_preds_flagged %>%
  summarise(
    n = n(),
    pct_any_extrapolation = round(mean(any_extrapolation, na.rm = TRUE) * 100, 2),
    pct_pm25_out = round(mean(out_pm25, na.rm = TRUE) * 100, 2),
    pct_temp_out = round(mean(out_temp, na.rm = TRUE) * 100, 2),
    pct_rh_out   = round(mean(out_rh, na.rm = TRUE) * 100, 2)
  )
```

Overall proportion of observations with one or more predictors outside the range seen in the co-located training data.
```{r, echo=FALSE, message=FALSE, warning=FALSE}

datatable(overall_extrap_summary, rownames = FALSE)
```

Monitor-level proportion of observations with one or more predictors outside the training range.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

monitor_extrap_summary <- modulair_preds_flagged %>%
  group_by(monitor) %>%
  summarise(
    n = n(),
    pct_any_extrapolation = mean(any_extrapolation, na.rm = TRUE) * 100,
    pct_pm25_out = mean(out_pm25, na.rm = TRUE) * 100,
    pct_temp_out = mean(out_temp, na.rm = TRUE) * 100,
    pct_rh_out   = mean(out_rh, na.rm = TRUE) * 100
  ) %>%
  mutate(across(starts_with("pct_"), ~ round(.x, 2))) %>%
  arrange(desc(pct_any_extrapolation))

datatable(monitor_extrap_summary, rownames = FALSE)
```

## Summarize calibration changes elastic net 

This section quantifies how much the calibrated predictions differ from the raw Modulair values.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
modulair_preds_flagged <- modulair_preds_flagged %>%
  mutate(
    delta_enet = pm25_fem_enet - mod_pm25,
    delta_xgb  = pm25_fem_xgb  - mod_pm25,
    ratio_enet = if_else(mod_pm25 > 0, pm25_fem_enet / mod_pm25, NA_real_),
    ratio_xgb  = if_else(mod_pm25 > 0, pm25_fem_xgb  / mod_pm25, NA_real_)
  )

calib_change_table <- modulair_preds_flagged %>%
  group_by(monitor) %>%
  summarise(
    n = n(),
    mean_raw = round(mean(mod_pm25, na.rm = TRUE),2),
    mean_enet = round(mean(pm25_fem_enet, na.rm = TRUE),2),
    mean_xgb  = round(mean(pm25_fem_xgb, na.rm = TRUE),2),
    mean_delta_enet = round(mean(delta_enet, na.rm = TRUE),2),
    mean_delta_xgb  = round(mean(delta_xgb, na.rm = TRUE),2),
    median_ratio_enet = median(ratio_enet, na.rm = TRUE),
    median_ratio_xgb  = median(ratio_xgb, na.rm = TRUE)
  ) %>%
  arrange(desc(abs(mean_delta_enet)))
```

Summary of average raw PM2.5, calibrated PM2.5, and calibration shifts by monitor.
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=14}

datatable(calib_change_table, rownames = FALSE)



ggplot(modulair_preds_flagged, aes(x = monitor, y = delta_enet)) +
  geom_boxplot() +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Monitor",
    y = "XGB calibration change (calibrated - raw)",
    title = "Calibration change by monitor"
  )
```

## calibration change versus meteorology

These plots show whether the size of the XGBoost calibration adjustment varies systematically with relative humidity or temperature.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(modulair_preds_flagged %>% 
  sample_n(10000), aes(x = mod_rh, y = delta_enet)) +
  geom_point(alpha = 0.2, size = 1) +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  labs(
    x = "Relative humidity",
    y = "XGB calibration change",
    title = "Calibration change vs RH"
  )


ggplot(modulair_preds_flagged %>% 
  sample_n(10000), aes(x = mod_temp, y = delta_enet)) +
  geom_point(alpha = 0.2, size = 1) +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  labs(
    x = "Temperature",
    y = "XGB calibration change",
    title = "Calibration change vs temperature"
  )

```



## Summarize calibration changes XG Boost 

This section quantifies how much the calibrated predictions differ from the raw Modulair values.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
modulair_preds_flagged <- modulair_preds_flagged %>%
  mutate(
    delta_enet = pm25_fem_enet - mod_pm25,
    delta_xgb  = pm25_fem_xgb  - mod_pm25,
    ratio_enet = if_else(mod_pm25 > 0, pm25_fem_enet / mod_pm25, NA_real_),
    ratio_xgb  = if_else(mod_pm25 > 0, pm25_fem_xgb  / mod_pm25, NA_real_)
  )

calib_change_table <- modulair_preds_flagged %>%
  group_by(monitor) %>%
  summarise(
    n = n(),
    mean_raw = round(mean(mod_pm25, na.rm = TRUE),2),
    mean_enet = round(mean(pm25_fem_enet, na.rm = TRUE),2),
    mean_xgb  = round(mean(pm25_fem_xgb, na.rm = TRUE),2),
    mean_delta_enet = round(mean(delta_enet, na.rm = TRUE),2),
    mean_delta_xgb  = round(mean(delta_xgb, na.rm = TRUE),2),
    median_ratio_enet = round(median(ratio_enet, na.rm = TRUE,2)),
    median_ratio_xgb  = round(median(ratio_xgb, na.rm = TRUE,2))
  ) %>%
  arrange(desc(abs(mean_delta_xgb)))
```

Summary of average raw PM2.5, calibrated PM2.5, and calibration shifts by monitor.
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=14}

datatable(calib_change_table, rownames = FALSE)



ggplot(modulair_preds_flagged, aes(x = monitor, y = delta_xgb)) +
  geom_boxplot() +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Monitor",
    y = "XGB calibration change (calibrated - raw)",
    title = "Calibration change by monitor"
  )
```

## calibration change versus meteorology

These plots show whether the size of the XGBoost calibration adjustment varies systematically with relative humidity or temperature.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(modulair_preds_flagged %>% 
  sample_n(10000), aes(x = mod_rh, y = delta_xgb)) +
  geom_point(alpha = 0.2, size = 1) +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  labs(
    x = "Relative humidity",
    y = "XGB calibration change",
    title = "Calibration change vs RH"
  )


ggplot(modulair_preds_flagged %>% 
  sample_n(10000), aes(x = mod_temp, y = delta_xgb)) +
  geom_point(alpha = 0.2, size = 1) +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  labs(
    x = "Temperature",
    y = "XGB calibration change",
    title = "Calibration change vs temperature"
  )

```

